---
title: "Analysis"
author: "Nicolas Banholzer"
date: "2024-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "..")
```

```{r libraries, include=FALSE}
library(tidyverse)
library(reshape2)
library(rstanarm)
library(brms)
library(rstan)
library(tidybayes)
library(mice)
library(klaR)
source("utils/plotting.R")
source("utils/summaries.r")
```


## Data

We analyze the quality of life (QoL) outcomes of tuberculosis (TB) patients 
at the start of, end of, and post treatment in IeDEA-SA cohorts:
South Africa, Malawi, Mozambique, Zambia, and Zimbabwe.
The  data is as of April 5, 2024.


```{r data, include=FALSE}
# data of patients with at least one outcome
qol <- readRDS("data-clean/phys-ment-data.rds") %>%
  arrange(time, record_id) %>%
  # cutoffs
  mutate(
    phq9_score_bin = ifelse(phq9_score > 10, 1, 0),
    sf12_phys_bin = ifelse(sf12_phys < 50, 1, 0),
    sf12_ment_bin = ifelse(sf12_ment < 42, 1, 0),
    smwt_dist_bin = ifelse(smwt_dist < 400, 1, 0),
    stst_nr_bin = ifelse(stst_nr < 20, 1, 0),
    sgrq_tot_score_bin = ifelse(sgrq_tot_score > 25, 1, 0),
    tb_symp_score_bin = ifelse(tb_symp_score > 4, 1, 0)
  )
qol_start <- filter(qol, time == "Start")
qol_end <- filter(qol, time == "End")
qol_post <- filter(qol, time == "Post")

# labeling
outcome_cont <- c(
  "phq9_score" = "PHQ9-S",
  "sf12_ment" = "SF12-MCS",
  "sf12_phys" = "SF12-PCS",
  "smwt_dist" = "6MWT-D",
  "stst_nr" = "STST-R",
  "sgrq_tot_score" = "SGRQ-TS",
  "tb_symp_score" = "TB-SS"
)

outcome_bin <- c(
  "phq9_score_bin" = "PHQ9-S>10",
  "sf12_ment_bin" = "SF12-MCS<42",
  "sf12_phys_bin" = "SF12-PCS<50",
  "smwt_dist_bin" = "6MWT-D<400m",
  "stst_nr_bin" = "STST-R<20",
  "sgrq_tot_score_bin" = "SGRQ-TS>25",
  "tb_symp_score_bin" = "TB-SS>4"
)

predictor <- c(
  "age" = "Age",
  "sex" = "Female",
  "hiv" = "HIV+",
  "highbact" = "High bact. load",
  "mdr" = "MDR",
  "cavity" = "Cavitation",
  "tbd_pat_cat" = "Relapse case",
  "as_alc_calc" = "Alcohol score",
  "as_tobacco_calc" = "Tobacco score"
)
predictor_bin <- c(
  "age_bin" = "Age<30",
  "sex" = "Female",
  "hiv" = "HIV+",
  "highbact" = "High bact. load",
  "mdr" = "MDR",
  "cavity" = "Cavitation",
  "tbd_pat_cat" = "Relapse case",
  "alc_bin" = "Drinker",
  "tobacco_bin" = "Smoker"
)

sites <- c("South Africa", "Malawi", "Mosambique", "Zambia", "Zimbabwe")

treat_effect_lab <- c("End vs Start treatment", "Post vs End treatment")
```

**Outcomes**:

The quality of life **(QoL)** outcomes:

* PHQ9-S: Patient Health Questionnaire 9 - Score (Range: 0-27)
* SF12-MCS: Short form survey (SF12) - Mental Component Score (Range: 0-100)
* SF12-PCS: Short form survey (SF12) - Physical Component Score (Range: 0-100)
* 6MWT-D: 6 Minute Walk Test - Distance
* STST-R: Sit-to-Stand Test - Number of Repetitions
* SGRQ-S: St. George's Respiratory Questionnaire - Total Score (Range: 0-100)
* TB-SS: TB - Symptom Score (Range: 0-9)

We define impaired QoL per outcome as:

1. PHQ9-S > 10
2. SF12-MCS < 42
3. SF12-PCS < 50
4. 6MWT-D < 400m
5. STST-R < 20
6. SGRQ-TS > 25
7. TB-SS > 4

**Exclusion:**

We exclude observations where we still wait for the next clinical visit.
That is, end of treatment or post treatment is still within 8mo since last visit.

**Missing observations**:

The outcomes can be missing for one of the following reasons:

1. *Death*: The patient died during or post treatment.
2. *Missing follow-up* (MFU): Missing QoL outcome because there was no follow-up visit.
3. *Unrecorded*: One or more QoL outcomes were not recorded.

**Patient characteristics**:

The predictors we consider are:

1. Age
2. Sex
3. HIV test result
4. High bacterial load on smear test.
5. Multi-drug resistant TB (MDR-TB)
6. Presence of cavitation on chest X-ray.
7. Relapse instead of new TB case.
8. ASSIST alcohol score
9. ASSIST tobacco score

**Additional variables::**

We further consider:

1. Any serious but non-fatal adverse event (AE) during treatment: 

* Life-threatining.
* Requires in-patient hospitalization or prolongation of existing hospitalization.
* Results in persistent or significant disability/incapacity.
* Results in congenital anomaly/birth defect.
* Other important medical event that may not be immediately life-threatening or result in death, or require hospitalisation, but may jeopardise the participant or may require intervention to prevent one of the other outcomes listed above.

2. Any mental health treatment (therapy or medication) since last visit:

* Mental health counselling, psychiatric medication, or hospital admission for mental illness.
* Treatment from a traditional healer for mental illness.
* Treatment from an outpatient health department (outpatient) for mental illness.
* Admission to a hospital or other health institution (inpatient) at least for one night for the treatment of mental illness
* Any medication for a mental illness.


**TB treatment outcome**:

The association with TB treatment outcome is currently not analyzed because it is mostly missing. 

## Descriptives

### Study population

```{r sample}
print(
  sprintf(
    "Total: %i patients with %i observations.",
    n_distinct(qol$record_id), nrow(qol)
  )
)
```

**Figure:** Number of observations by time.

```{r observations}
qol %>%
  group_by(time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x = time, y = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), size = 8 / cm(1), vjust = -1) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(y = "No. of observations") +
  theme_custom() +
  theme(
    axis.title.x = element_blank()
  )
```

**Table**: Proportion of patients with missing follow-up (MFU).

```{r lost-to-follow-up}
qol %>%
  group_by(time) %>%
  summarize(
    `no. of MFU` = sum(mfu),
    `proportion` = paste0(round(sum(mfu) / n() * 100), "%")
  ) %>%
  knitr::kable(align = "lrr")
```

**Table**: Proportion of patients who died.

```{r deaths}
qol %>%
  group_by(time) %>%
  summarize(
    `no. of deaths` = sum(death),
    `proportion` = paste0(round(sum(death) / n() * 100), "%")
  ) %>%
  knitr::kable(align = "lrr")
```

**Table**: Patient characteristics at baseline.

```{r predictors}
qol_start %>%
  dplyr::select(all_of(names(predictor))) %>%
  gather() %>%
  mutate(
    key = recode(key, !!!predictor),
    key = factor(key, levels = predictor)
  ) %>%
  group_by(key) %>%
  summarize_all(
    ~ ifelse(max(.x, na.rm = TRUE) > 1, median_iqr(.x), count_bin(.x))
  ) %>%
  ungroup() %>%
  arrange(key) %>%
  set_names(c("Variable", "Summary")) %>%
  knitr::kable(align = "lr")
```

**Table**: Distribution of patient characteristics by time.

```{r}
qol %>%
  filter(!mfu, !death) %>%
  mutate(
    age_bin = ifelse(age < 30, 1, 0),
    alc_bin = ifelse(as_alc_calc > 0, 1, 0),
    tobacco_bin = ifelse(as_tobacco_calc > 0, 1, 0)
  ) %>%
  dplyr::select(all_of(c("time", names(predictor_bin)))) %>%
  melt("time") %>%
  mutate(variable = recode(variable, !!!predictor_bin)) %>%
  group_by(time, variable) %>%
  summarise_all(
    ~ ifelse(max(.x, na.rm = TRUE) > 1, median_iqr(.x), count_bin(.x))
  ) %>%
  ungroup() %>%
  pivot_wider(names_from = time, values_from = value) %>%
  knitr::kable(align = "lrrr")
```

Note: 

- Almost no follow-up visits among MDR patients.
- Lower prevalence of alcohol and tobacco use at the end of treatment (but not post treatment) compared with baseline.


**Table**: Share of sites.

```{r sites}
qol_start %>%
  group_by(site) %>%
  summarize(`no. of sites` = n()) %>%
  ungroup() %>%
  mutate(proportion = round(`no. of sites` / sum(`no. of sites`) * 100)) %>%
  knitr::kable(align = "lrr")
```

**Table**: Number of patients with non-fatal serious adverse events (AE). 

```{r no-adverse-events}
qol %>%
  group_by(time) %>%
  summarize(
    n_tot = sum(nf_ae == 1, na.rm = TRUE),
    n_only = sum(nf_ae == 1 & !death & !mfu, na.rm = TRUE),
    n_died = sum(nf_ae == 1 & death, na.rm = TRUE),
    n_mfu = sum(nf_ae == 1 & mfu, na.rm = TRUE)
  ) %>%
  set_names(c("Time", "AE", "Only AE", "AE and death", "AE and MFU")) %>%
  knitr::kable(align = "lrrrr")
```

Note:

- The actual number of adverse events is 29 but 21 are from patients where we still wait for the end of treatment visit.
- Patients may have a non-fatal AE but still die (?)

**Table**: List of visits of patients with any AE.

```{r list-adverse-events}
qol %>%
  group_by(record_id) %>%
  filter(any(nf_ae == 1)) %>%
  ungroup() %>%
  dplyr::select(record_id, time, date_visit, nf_ae_date) %>%
  arrange(record_id, time) %>%
  set_names(c("Record ID", "Time", "Visit date", "AE date")) %>%
  knitr::kable(align = "llrr")
```

**Table**: Number of patients who received mental health treatments. 

```{r no-mental-health-treatment}
qol %>%
  group_by(time) %>%
  summarize(
    `no. of patients` = sum(ment_trt_yn == 1, na.rm = TRUE),
    `proportion` = paste0(round(sum(ment_trt_yn == 1, na.rm = TRUE) / n() * 100), "%")
  ) %>%
  knitr::kable(align = "lrr")
```

**Table**: List of visits of patients receiving mental health treatment.

```{r list-mental-health-treatment}
qol %>%
  group_by(record_id) %>%
  filter(any(ment_trt_yn == 1)) %>%
  ungroup() %>%
  dplyr::select(record_id, time, phq9_score, sf12_ment, death, mfu, ment_trt_yn) %>%
  mutate(
    death = ifelse(death, "Yes", "No"),
    mfu = ifelse(mfu, "Yes", "No"),
    ment_trt_yn = ifelse(ment_trt_yn == 1, "Yes", "No")
  ) %>%
  mutate(sf12_ment = round(sf12_ment)) %>%
  arrange(record_id, time) %>%
  set_names(c("Record ID", "Time", "PHQ9-S", "SF12-MCS", "Death", "MFU", "Mental health treat.")) %>%
  knitr::kable(align = "llrrrr")
```

**Table**: TB Treatment outcomes.

```{r tb-treatment-outcomes}
qol %>%
  filter(time == "End") %>%
  mutate(
    `treatment outcome` = ifelse(is.na(treat_success), "Not available",
      ifelse(treat_success == 1, "Success", "Failure")
    ),
    `treatment outcome` = factor(`treatment outcome`,
      levels = c("Success", "Failure", "Not available")
    )
  ) %>%
  group_by(`treatment outcome`) %>%
  summarize(`no. of sites` = n()) %>%
  ungroup() %>%
  mutate(proportion = round(`no. of sites` / sum(`no. of sites`) * 100)) %>%
  knitr::kable(align = "lrr")
```

### Outcomes

#### Summary

**Table**: Summary of continuous QoL outcomes.

```{r summaries-1}
# start of treatment
qol_start %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(
    key = recode(key, !!!outcome_cont),
    key = factor(key, levels = outcome_cont)
  ) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment start")) %>%
  knitr::kable(align = "lr")
```

```{r summaries-2}
# end of treatment
qol_end %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(
    key = recode(key, !!!outcome_cont),
    key = factor(key, levels = outcome_cont)
  ) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment end")) %>%
  knitr::kable(align = "lr")
```

```{r summaries-3}
# post treatment
qol_post %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(
    key = recode(key, !!!outcome_cont),
    key = factor(key, levels = outcome_cont)
  ) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) post treatment")) %>%
  knitr::kable(align = "lr")
```

#### Distribution

**Figure**: Boxplots of continuous QoL outcomes by time.

```{r qol-outcomes-boxplot-by-time}
qol_boxplot <- qol %>%
  dplyr::select(all_of(c("time", names(outcome_cont)))) %>%
  melt("time") %>%
  rename(outcome = variable) %>%
  mutate(outcome = recode(outcome, !!!outcome_cont)) %>%
  filter(!is.na(value)) %>%
  mutate(
    outcome_cat = ifelse(
      outcome %in% c("PHQ9-S", "SF12-MCS"),
      "Mental health outcomes",
      ifelse(outcome %in% c("SF12-PCS", "6MWT-D", "STST-R"),
        "Physical health outcomes", "TB symptoms scores"
      )
    ),
    outcome_cat = factor(outcome_cat, levels = c(
      "Mental health outcomes", "Physical health outcomes", "TB symptoms scores"
    ))
  ) %>%
  ggplot(aes(y = value, x = time, fill = time)) +
  facet_wrap(~outcome, scales = "free_y", ncol = 3) +
  geom_boxplot() +
  labs(y = "Score value") +
  scale_fill_brewer(palette = "Set2") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.title = element_blank()
  )

save_plot(
  qol_boxplot,
  pdf_file = "results/outcome-boxplot.png",
  w = 16, h = 16
)
```

**Figure**: Histograms of continuous QoL outcomes by time.

```{r histogram-by-time, fig.width=6.3, fig.height=8.7}
qol_denesity <- qol %>%
  dplyr::select(all_of(c("time", names(outcome_cont)))) %>%
  melt("time") %>%
  rename(outcome = variable) %>%
  mutate(outcome = recode(outcome, !!!outcome_cont)) %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x = value, group = time)) +
  ggh4x::facet_grid2(outcome ~ time, scales = "free", independent = "x") +
  geom_histogram(aes(y = ..density..)) +
  coord_cartesian(expand = FALSE) +
  labs(y = "Density") +
  theme_custom() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

save_plot(
  qol_denesity,
  pdf_file = "results/outcome-density-distribution.png",
  w = 16, h = 22
)
```

**Figure**: Pairwise correlations between QoL outcomes.

```{r pairwise-correlation, fig.width=10, fig.height=10}
my_cor <- function(data, mapping, ...) {
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  cor_val <- cor(x, y, use = "pairwise.complete.obs")
  ggally_text(paste0("r = ", round(cor_val, 2)), size = 8 / cm(1), color = "black", ...)
}

my_points <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(shape = 1, alpha = .5, ...)
}

qol %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  set_names(outcome_cont) %>%
  GGally::ggpairs(
    lower = list(continuous = my_points),
    upper = list(continuous = my_cor)
  ) +
  theme_custom()
```

**Figure:** Proportion of patients with impaired QoL.

```{r proportions, fig.width=6.3, fig.height=3.9}
var_names <- c(
  "Y" = "Yes (impaired QoL)",
  "N" = "No (QoL not impaired)",
  "D" = "Death before visit",
  "L" = "Missing follow-up",
  "U" = "Unrecorded"
)

var_colors <- c(
  wes_palette("Zissou1")[5],
  wes_palette("Zissou1")[2],
  wes_palette("Zissou1")[3],
  wes_palette("Royal1")[3],
  wes_palette("Royal1")[1]
)

qol_bin_comp <- qol %>%
  dplyr::select(all_of(c("time", "mfu", "death", names(outcome_bin)))) %>%
  melt(c("time", "mfu", "death")) %>%
  group_by(time, variable) %>%
  summarize(
    Y = sum(value, na.rm = TRUE),
    N = sum(value == 0, na.rm = TRUE),
    D = sum(death),
    U = sum(is.na(value) & !mfu & !death, na.rm = TRUE),
    L = sum(mfu, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  rename(outcome = variable) %>%
  melt(c("time", "outcome")) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin),
    variable = recode(variable, !!!var_names),
    variable = factor(variable, levels = var_names)
  ) %>%
  group_by(time, outcome) %>%
  mutate(p = value / sum(value) * 100) %>%
  ungroup()

qol_bin_comp_pl <- ggplot(data = qol_bin_comp, mapping = aes(x = time, y = p)) +
  facet_wrap(~outcome, nrow = 1) +
  geom_bar(
    aes(fill = variable),
    stat = "identity",
    position = position_stack(reverse = TRUE)
  ) +
  geom_text(
    data = filter(qol_bin_comp, variable == "Yes (impaired QoL)"),
    mapping = aes(group = time, label = paste0(round(p))),
    vjust = -.5, size = 6 / cm(1),
    color = var_colors[1],
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = var_colors,
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  labs(
    y = "Proportion of patients (%)",
    x = "Treatment",
    fill = "a"
  ) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.title = element_blank(),
    legend.key.width = unit(.2, "cm"),
    legend.key.height = unit(.4, "cm")
  )

save_plot(
  qol_bin_comp_pl,
  pdf_file = "results/impaired-qol.png",
  w = 16, h = 10
)
```

#### Patient trajectories for depression

**Figure**: Trajectories of patients who were depressed (defined as PHQ9-S > 10) once during the study.

```{r depressed-flow, fig.width=6.3, fig.height=3.9}
var_names_dprs <- c(
  "Y" = "Depressed",
  "N" = "Not depressed",
  "D" = "Death before visit",
  "L" = "Missing follow-up",
  "U" = "Unrecorded"
)

var_colors_dprs <- c(
  wes_palette("Zissou1")[5],
  wes_palette("Zissou1")[2],
  wes_palette("Zissou1")[3],
  wes_palette("Royal1")[3],
  wes_palette("Royal1")[1]
)

# create links
links_dprs <- qol %>%
  dplyr::select(
    time,
    record_id,
    ment_trt_yn,
    death,
    mfu,
    phq9_score_bin
  ) %>%
  rename(value = phq9_score_bin) %>%
  mutate(
    value =
      ifelse(death == 1 & is.na(value) & time != "Start", var_names_dprs["D"],
        ifelse(mfu, var_names_dprs["L"],
          ifelse(is.na(value), var_names_dprs["U"],
            ifelse(value == 1, var_names_dprs["Y"], var_names_dprs["N"])
          )
        )
      ),
    value = factor(value, levels = var_names_dprs),
    ment_trt_yn = ifelse(ment_trt_yn == 1, "Yes", "No"),
    ment_trt_yn = factor(ment_trt_yn, levels = c("No", "Yes"))
  ) %>%
  # filter patients that were depressed at least once
  group_by(record_id) %>%
  filter(any(value == var_names_dprs["Y"])) %>%
  ungroup() %>%
  group_by(time, value) %>%
  mutate(freq = n()) %>%
  ungroup()

flow_dprs_pl <- links_dprs %>%
  ggplot(aes(
    x = time,
    stratum = value,
    alluvium = record_id,
    fill = value,
    label = freq
  )) +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft") +
  geom_stratum() +
  geom_text(stat = "stratum", size = 8 / cm(1)) +
  scale_fill_manual(values = var_colors_dprs) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) +
  scale_x_discrete(labels = c("Start", "End", "Post")) +
  labs(x = "Treatment") +
  theme_custom() +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    axis.title.y = element_blank()
  )

save_plot(
  flow_dprs_pl,
  pdf_file = "results/sankey-depressed.png",
  w = 16, h = 10
)
```

Remark:

- Note that none of the depressed patients received any mental health treatment during TB treatment.

## Imputation

We impute the missing values in the continuous QoL outcomes.
Outcomes missing because of death or missing follow-up visit will not be imputed, only unrecorded ones.
However, we consider subsequent death or missing follow-up visit as a predictor during imputation.
Similarly, we consider any adverse adverse event at the start of, end of, or post treatment as a predictor during imputation.

**Figure**: Imputation check for a random sample of 30 patients, showing the imputed majority label across datasets.

```{r imputation, fig.height=12}
# wide format
qol_fu <- qol %>%
  group_by(record_id) %>%
  arrange(time) %>%
  mutate(mfu_lead = lead(mfu), death_lead = lead(death)) %>%
  ungroup() %>%
  filter(!mfu & !death) %>%
  ungroup() %>%
  arrange(record_id, time) %>%
  mutate(
    across(
      c(
        mfu_lead, death_lead, ment_trt_yn,
        sex, hiv, mdr, cavity, tbd_pat_cat
      ),
      ~ factor(ifelse(.x == 1, "Yes", "No"), levels = c("No", "Yes"))
    )
  ) %>%
  dplyr::select(all_of(c(
    "record_id",
    "time",
    "site",
    "mfu_lead",
    "death_lead",
    "nf_ae",
    "ment_trt_yn",
    names(outcome_cont),
    names(predictor)
  )))

# imputation
n_datasets <- 10
qol_fu_imp <- mice(
  qol_fu[, -1],
  m = n_datasets,
  maxit = 10,
  seed = 12345,
  print = FALSE
)

# plot(qol_fu_imp, layout = c(2, 16))

# data
qol_fu_imp <- complete(qol_fu_imp, "long") %>%
  mutate(record_id = rep(qol_fu$record_id, times = n_datasets))

qol_fu_imp_mode <- qol_fu_imp %>%
  dplyr::select(all_of(c(".imp", "record_id", "time", names(outcome_cont)))) %>%
  melt(c(".imp", "record_id", "time")) %>%
  group_by(record_id, time, variable) %>%
  summarize(
    value = mean(value)
  ) %>%
  rename(imputed = value) %>%
  left_join(
    qol_fu %>%
      dplyr::select(all_of(c("record_id", "time", names(outcome_cont)))) %>%
      melt(c("record_id", "time")) %>%
      rename(observed = value),
    by = c("record_id", "time", "variable")
  ) %>%
  mutate(is_observed = ifelse(is.na(observed), FALSE, TRUE)) %>%
  rename(outcome = variable)

check_imp_pl <- qol_fu_imp_mode %>%
  group_by(record_id) %>%
  filter(any(!is_observed)) %>%
  ungroup() %>%
  filter(record_id %in% sample(unique(record_id), 30)) %>%
  ggplot(aes(
    y = time,
    x = outcome,
    color = is_observed,
    label = round(imputed)
  )) +
  geom_text(size = 8 / cm(1)) +
  facet_wrap(~record_id, ncol = 3) +
  labs(color = "Is observed") +
  scale_x_discrete(labels = function(x) outcome_cont) +
  scale_y_discrete(labels = function(x) c("Start", "End", "Post")) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

save_plot(
  check_imp_pl,
  pdf_file = "results/imputation-check.png",
  w = 16, h = 30
)
```


## Modeling

We model QoL as a multivariate outcome to consider correlation between the outcomes.
We model the log outcomes to make the results more interpretable and comparable.
Estimates differences in QoL over time are reported as the percent changes in QoL outcomes. 
Estimates associations are reported as percent change in QoL for a one-unit change in the predictor.
We report the median and 95% credible interval.

```{r model}
# prepare data
qol_mdat <- qol_fu_imp %>%
  dplyr::select(all_of(c(
    ".imp",
    "record_id",
    "time",
    "site",
    "mfu_lead",
    "death_lead",
    names(outcome_cont),
    names(predictor)
  ))) %>%
  mutate(
    alc_bin = ifelse(as_alc_calc > 0, "Yes", "No"),
    tobacco_bin = ifelse(as_tobacco_calc > 0, "Yes", "No"),
    age_bin = ifelse(age < 30, "Yes", "No"),
    highbact = ifelse(highbact == 1, "Yes", "No"),
    across(
      c("mfu_lead", "death_lead", names(predictor_bin)),
      ~ ifelse(.x == "Yes", 1, 0)
    )
  ) %>%
  dplyr::select(-age) %>%
  mutate(
    across(c(all_of(names(outcome_cont))), log1p)
  ) %>%
  mutate(dummy = 1) %>%
  pivot_wider(names_from = site, values_from = dummy, values_fill = 0) %>%
  mutate(dummy = 1) %>%
  pivot_wider(names_from = time, values_from = dummy, values_fill = 0) %>%
  mutate(
    across(
      c(names(predictor_bin), "as_alc_calc", "as_tobacco_calc"),
      ~ .x * End,
      .names = "{.col}_end"
    ),
    across(
      c(names(predictor_bin), "as_alc_calc", "as_tobacco_calc"),
      ~ .x * Post,
      .names = "{.col}_post"
    )
  ) %>%
  dplyr::select(
    -`South Africa`,
    -`Start`,
    -mdr_end,
    -mdr_post
  )

for (i in unique(qol_mdat$.imp)) {
  qol_mdat %>%
    filter(.imp == i) %>%
    dplyr::select(-.imp, -record_id) %>%
    write.csv(
      file = paste0("fitted-models/model-data/imp-data-", i, ".csv"),
      row.names = FALSE
    )
}

# ... run models in numpyro ... #

# collect posterior samples
mod_smpls <- data.frame(
  file = list.files("fitted-models/model-samples", full.names = TRUE)
) %>%
  mutate(
    imp = as.numeric(stringi::stri_extract(file, regex = "\\d")),
    data = lapply(file, read.csv)
  ) %>%
  unnest(data)
```

### QoL changes over time

```{r modeled-qol-changes}
# plot
treat_effect_pl <- mod_smpls %>%
  filter(predictor %in% c("End", "Post")) %>%
  mutate(
    predictor = recode(predictor, End = "End vs Start", Post = "Post vs Start"),
    predictor = factor(predictor, levels = c("End vs Start", "Post vs Start")),
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  ) %>%
  group_by(outcome, predictor) %>%
  summarize(
    median = median(beta),
    lower = quantile(beta, 0.025),
    upper = quantile(beta, 0.975)
  ) %>%
  ggplot(aes(x = median, y = outcome, color = predictor)) +
  geom_vline(aes(xintercept = 0), linetype = "dotted", color = "red") +
  geom_point(position = position_dodge(width = .66)) +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    position = position_dodge(width = .66),
    height = .25
  ) +
  labs(x = "Estimated change (%)") +
  scale_x_continuous(labels = scales::percent_format(suffix = "")) +
  scale_color_manual(values = RColorBrewer::brewer.pal(name = "Set2", n = 3)[-1]) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank()
  )

save_plot(
  treat_effect_pl,
  pdf_file = "results/qol-changes.png",
  w = 8, h = 8
)
```

Interpretation:

- All QoL outcomes improved from the start to the end of treatment.
- Improvements post treatment were small.


### Sensitivity check for death and MFU patients

In the following, we estimate for each outcome the change in QoL that would be needed 
in patients who died or with missing follow-up visits 
to offset the estimated QoL changes in patients with follow-up visits.
Basically, what we compute is *-beta x (n_fu / n_nfu)* for each outcome,
where n_fu is the number of follow-up patients and n_nfu is the number of patients with missing follow-up or death.

**Figure:** Estimated change in QoL (grey) vs offsetting hypothetical change in QoL for patients with missing follow-up visits or death (red).

```{r sensitivity-check}
n_end <- n_distinct(qol$record_id[qol$time == "End"])
n_end_mfu <- sum(qol$mfu[qol$time == "End"])
n_end_death <- sum(qol$death[qol$time == "End"])
n_end_fu <- n_end - n_end_mfu - n_end_death
scale_fct <- 1

null_treat_effect_pl <- mod_smpls %>%
  filter(predictor == "End") %>%
  mutate(
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  ) %>%
  mutate(
    n_nfu = n_end_mfu + n_end_death,
    n_fu = n_end_fu,
    beta_zero = -(n_fu / n_nfu) * beta
  ) %>%
  dplyr::select(outcome, beta_zero, beta) %>%
  melt("outcome") %>%
  mutate(
    group = ifelse(variable == "beta",
      "Follow-up patients (estimated)",
      "Missing follow-up/death (hypothetical)"
    ),
    group = factor(group, levels = c(
      "Follow-up patients (estimated)",
      "Missing follow-up/death (hypothetical)"
    ))
  ) %>%
  ggplot(aes(x = value, y = outcome, fill = group)) +
  geom_vline(aes(xintercept = 0), linetype = "dotted", color = "red") +
  stat_halfeye() +
  scale_x_continuous(labels = scales::percent_format(suffix = "")) +
  scale_fill_manual(values = wes_palette("Royal1")) +
  labs(x = "Estimated change (%)") +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank()
  )

null_treat_effect_pl
```

Interpretation:

- The deterioration in QoL for patients with missing follow-up visits or death would need to be extremely large 
to offset the improvements at the end of treatment.


### Association with patient characteristics

We define significant associations as those with a 95% credible interval (CrI) not overlapping zero.
Since we report many associations, significant ones have to be interpreted with caution.
The reason is that with some many comparisons, we should expect some false positives.
(We could remedy this using shrinkage priors such as the horeshoe prior. 
These priors would shrink most estimated associations towards zero, except for the ones that appear strong.)

The following predictors are transformed:

- Age: dichotomized into age < 30 and age >= 30.
- Drinker and Smoker: dichotomized into any alcohol or tobacco use denoted as an ASSIST score greater than zero.
- The numeric ASSIST alcohol and tobacco score are still also used as continuous predictors in the model.

The following variables are also included:

- Subsequent MFU: To check whether missing follow-up is associated with QoL at baseline.
- Subsequent death: To check whether subsequent death is associated with QoL at baseline.

We include time interaction effects for all predictors to estimate the association with baseline, end of treatment, and post treatment QoL. 

**Figure**: Estimated association of QoL outcomes with patient characteristics. 

```{r association-by-outcome, fig.height=12, fig.width=10}
# plot associations of binary predictors
assoc <- rbind(
  mod_smpls %>%
    filter(predictor %in% names(predictor_bin)) %>%
    mutate(time = "Start"),
  mod_smpls %>%
    filter(predictor %in% paste0(names(predictor_bin), "_end")) %>%
    mutate(predictor = gsub("_end", "", predictor)) %>%
    mutate(time = "End"),
  mod_smpls %>%
    filter(predictor %in% paste0(names(predictor_bin), "_post")) %>%
    mutate(predictor = gsub("_post", "", predictor)) %>%
    mutate(time = "Post")
) %>%
  pivot_wider(values_from = beta, names_from = time) %>%
  mutate(End = Start + End, Post = Start + Post) %>%
  pivot_longer(
    cols = c("Start", "End", "Post"),
    names_to = "time",
    values_to = "beta"
  ) %>%
  mutate(
    predictor = recode(predictor, !!!predictor_bin),
    predictor = factor(predictor, levels = predictor_bin),
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont),
    time = factor(time, levels = c("Start", "End", "Post"))
  ) %>%
  filter(!(predictor == "MDR" & time %in% c("End", "Post")))

assoc_pl <- assoc %>%
  group_by(outcome, predictor, time) %>%
  summarize(
    median = median(beta),
    lower = quantile(beta, 0.025),
    upper = quantile(beta, 0.975)
  ) %>%
  ungroup() %>%
  ggplot(aes(x = median, y = outcome, color = time)) +
  facet_wrap(~predictor, ncol = 3) +
  geom_vline(aes(xintercept = 0), linetype = "dotted", color = "red") +
  geom_point(position = position_dodge(width = .66)) +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    position = position_dodge(width = .66),
    height = .25
  ) +
  labs(x = "Estimated change (%)") +
  scale_x_continuous(labels = scales::percent_format(suffix = "")) +
  scale_color_brewer(palette = "Set2") +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank()
  )

save_plot(
  assoc_pl,
  pdf_file = "results/association.png",
  w = 16, h = 21
)
```

Interpretation: 

- Female sex was overall associated with worse QoL outcomes, especially at baseline and end of treatment.
- HIV+ was associated with worse PHQ9 score at baseline and end of treatment.
- High bacterial load and cavitation were associated with worse symptom scores at baseline but better scores at the end of treatment.
- Relapse cases were overall associated with worse QoL outcomes at all time points, especially symptom scores at baseline and end of treatment.
- Smoking tended to be overall associated with worse QoL outcomes. 
- MDR was only considered at baseline because there are only few MDR patients with follow-up visits.


**Figure** Probability that the predictors are associated with worse QoL outcomes overall.

```{r assoc-prob-overall}
assoc_overall_pl <- assoc %>%
  mutate(beta = ifelse(
    outcome %in% c("PHQ9-S", "SGRQ-TS", "TB-SS"),
    -beta, beta
  )) %>%
  group_by(time, predictor) %>%
  summarise(prob = mean(beta < 0)) %>%
  ungroup() %>%
  ggplot(aes(x = prob, y = predictor, fill = time)) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(width = 1, padding = .1)
  ) +
  geom_text(
    aes(label = scales::percent(prob, accuracy = 1)),
    position = position_dodge2(width = 1, padding = .1),
    hjust = 1,
    size = 8 / cm(1)
  ) +
  labs(x = "Probability of association with worse QoL (%)") +
  scale_x_continuous(
    labels = function(x) x * 100,
    limits = c(0, 1),
    expand = c(0, 0)
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank()
  )

save_plot(
  assoc_overall_pl,
  pdf_file = "results/association-prob-overall.png",
  w = 16, h = 16
)
```

Interpretation:

- Smoking most likely associated with worse QoL outcomes at all time points.
- Relapse cases most likely associated with worse QoL outcomes at the end of treatment.
- HIV+ and cavitation most likely associated with worse QoL outcomes at baseline.
- Age<30 most likely associated with better QoL outcomes at the end of treatment (just consider the inverse probability).


**Figure:** Probability that the predictors are associated with worse QoL outcomes by category.

```{r assoc-prob-category}
assoc_cat_pl <- assoc %>%
  mutate(
    beta = ifelse(
      outcome %in% c("PHQ9-S", "SGRQ-TS", "TB-SS"),
      -beta, beta
    ),
    category = ifelse(outcome %in% c("PHQ9-S", "SF12-MCS"),
      "Mental health scores", ifelse(outcome %in% c("SGRQ-TS", "TB-SS"),
        "TB symptom scores", "Physicial health scores"
      )
    ),
    category = factor(category, levels = c(
      "Mental health scores", "Physicial health scores", "TB symptom scores"
    ))
  ) %>%
  group_by(category, time, predictor) %>%
  summarise(prob = mean(beta < 0)) %>%
  ungroup() %>%
  ggplot(aes(x = prob, y = predictor, fill = time)) +
  facet_wrap(~category, ncol = 3) +
  geom_bar(
    stat = "identity",
    position = position_dodge2(width = 1, padding = .1)
  ) +
  geom_text(
    aes(label = scales::percent(prob, accuracy = 1, suffix = "")),
    position = position_dodge2(width = 1, padding = .1),
    hjust = -.5,
    size = 6 / cm(1)
  ) +
  labs(x = "Probability of association with worse QoL (%)") +
  scale_x_continuous(
    labels = function(x) x * 100,
    limits = c(0, 1),
    breaks = seq(0, 1, .25),
    expand = expansion(add = c(0, 0.175))
  ) +
  scale_fill_brewer(palette = "Set2") +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank()
  )

save_plot(
  assoc_cat_pl,
  pdf_file = "results/association-prob-by-category.png",
  w = 16, h = 16
)
```

Interpretation:

- The associations of relapse case and smoking are strongest in mental health and TB symptom scores.
- The associations age<30 are strongest in physical health scores.


**Table**: Estimated change in QoL for a one-unit increase in ASSIST alcohol and tobacco score.

```{r assoc-assist-score}
predictor_cont <- predictor[c("as_alc_calc", "as_tobacco_calc")]
assoc_cont <- rbind(
  mod_smpls %>%
    filter(predictor %in% names(predictor_cont)) %>%
    mutate(time = "Start"),
  mod_smpls %>%
    filter(predictor %in% paste0(names(predictor_cont), "_end")) %>%
    mutate(predictor = gsub("_end", "", predictor)) %>%
    mutate(time = "End"),
  mod_smpls %>%
    filter(predictor %in% paste0(names(predictor_cont), "_post")) %>%
    mutate(predictor = gsub("_post", "", predictor)) %>%
    mutate(time = "Post")
) %>%
  pivot_wider(values_from = beta, names_from = time) %>%
  mutate(End = Start + End, Post = Start + Post) %>%
  pivot_longer(
    cols = c("Start", "End", "Post"),
    names_to = "time",
    values_to = "beta"
  ) %>%
  mutate(
    predictor = recode(predictor, !!!predictor_cont),
    predictor = factor(predictor, levels = predictor_cont),
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont),
    time = factor(time, levels = c("Start", "End", "Post"))
  )

assoc_cont %>%
  group_by(outcome, predictor, time) %>%
  summarize(
    median = median(beta),
    lower = quantile(beta, 0.025),
    upper = quantile(beta, 0.975)
  ) %>%
  ungroup() %>%
  ggplot(aes(x = median, y = outcome, color = time)) +
  facet_wrap(~predictor, ncol = 3) +
  geom_vline(aes(xintercept = 0), linetype = "dotted", color = "red") +
  geom_point(position = position_dodge(width = .66)) +
  geom_errorbarh(
    aes(xmin = lower, xmax = upper),
    position = position_dodge(width = .66),
    height = .25
  ) +
  labs(x = "Estimated change (%)") +
  scale_x_continuous(labels = scales::percent_format(suffix = "")) +
  scale_color_brewer(palette = "Set2") +
  theme_custom() +
  theme(axis.title.y = element_blank())
```

Intepretation: 

- ASSIST alcohol score was associated with worse PHQ9 score post treatment.


**Table**: Estimated association of subsequent death with QoL.

```{r assoc-subsequent-death}
mod_smpls %>%
  filter(predictor == "death_lead") %>%
  mutate(
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  ) %>%
  group_by(outcome) %>%
  summarize(
    median = median(beta),
    lower = quantile(beta, 0.025),
    upper = quantile(beta, 0.975)
  ) %>%
  mutate_if(is.numeric, scales::percent, accuracy = 0.01) %>%
  knitr::kable(align = "lrrr")
```

Intepretation:

- Subsequent death was associated with worse QoL outcomes, except for TB-SS.


**Table**: Estimated association of subsequent missing follow-up with QoL at baseline.

```{r assoc-subsequent-mfu}
mod_smpls %>%
  filter(predictor == "mfu_lead") %>%
  mutate(
    outcome = recode(outcome, !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  ) %>%
  group_by(outcome) %>%
  summarize(
    median = median(beta),
    lower = quantile(beta, 0.025),
    upper = quantile(beta, 0.975)
  ) %>%
  mutate_if(is.numeric, scales::percent, accuracy = 0.01) %>%
  knitr::kable(align = "lrrr")
```

Intepretation:

- Subsequent missing follow-up was only significantly associated with 6MWT-D.


### QoL impairement over time

**Figure**: Estimated proportion of patients with impaired QoL (Pr = Bayesian probability that the proportions are unequal).

```{r model-qol-impairment, results = 'hide'}
# prepare data
qol_imp_bin <- qol_fu_imp %>%
  dplyr::select(
    .imp,
    record_id,
    time,
    names(outcome_cont)
  ) %>%
  mutate(
    phq9_score_bin = ifelse(phq9_score > 10, 1, 0),
    sgrq_tot_score_bin = ifelse(sgrq_tot_score > 25, 1, 0),
    tb_symp_score_bin = ifelse(tb_symp_score > 4, 1, 0),
    smwt_dist_bin = ifelse(smwt_dist < 400, 1, 0),
    stst_nr_bin = ifelse(stst_nr < 20, 1, 0),
    sf12_phys_bin = ifelse(sf12_phys < 50, 1, 0),
    sf12_ment_bin = ifelse(sf12_ment < 42, 1, 0)
  ) %>%
  group_by(.imp, time) %>%
  summarize(
    across(names(outcome_bin), sum),
    N = n()
  ) %>%
  melt(c(".imp", "time", "N")) %>%
  ungroup() %>%
  rename(n = value) %>%
  mutate(
    end = ifelse(time %in% c("End", "Post"), 1, 0),
    post = ifelse(time == "Post", 1, 0)
  ) %>%
  dplyr::select(-time) %>%
  nest(data = -.imp)

# run models
qol_bin_smpls <- tibble(
  outcome = c(),
  end_v_start = c(),
  post_v_end = c(),
)
for (out in names(outcome_bin)) {
  mod_out <- brm_multiple(
    formula = bf(
      n | trials(N) ~ end + post,
      family = binomial
    ),
    data = lapply(qol_imp_bin$data, function(x) filter(x, variable == out)),
    iter = 2000,
    warmup = 1000,
    chains = 1,
    cores = 1,
    seed = 12345,
    file = paste0("fitted-models/binary-models/", out, ".rds")
  )
  qol_bin_smpls_out <- tibble(
    outcome = out,
    end_v_start = unlist(posterior_samples(mod_out, "b_end")),
    post_v_end = unlist(posterior_samples(mod_out, "b_post"))
  )
  qol_bin_smpls <- rbind(qol_bin_smpls, qol_bin_smpls_out)
}
```

```{r qol-impairment-over-time, fig.width=7.9, fig.height=4}
# Bayesian posterior probability
qol_bin_pp <- qol_bin_smpls %>%
  group_by(outcome) %>%
  summarize(
    Pr_end = 100 * mean(end_v_start < 0),
    Pr_post = 100 * mean(post_v_end < 0)
  ) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin)
  ) %>%
  melt("outcome") %>%
  mutate(
    variable = recode(variable, Pr_end = "End", Pr_post = "Post"),
    variable = factor(
      variable,
      levels = c("End", "Post")
    )
  ) %>%
  mutate(
    value = scales::percent(value / 100, accuracy = 1),
    value = ifelse(value == "100%", "Pr>99%", paste0("Pr" = value))
  ) %>%
  rename(time = variable) %>%
  arrange(outcome, time) %>%
  pull(value)


# proportion of impairments
qol_imp_prop <- qol_imp_bin %>%
  unnest(data) %>%
  mutate(p = 100 * n / N) %>%
  mutate(
    time = ifelse(end == 0, "Start", ifelse(post == 1, "Post", "End")),
    time = factor(time, levels = c("Start", "End", "Post"))
  ) %>%
  rename(outcome = variable) %>%
  group_by(outcome, time) %>%
  summarise(
    p_m = mean(p),
    p_s = sd(p)
  ) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin)
  ) %>%
  arrange(outcome, time)

pval_ypos_end <- qol_imp_prop %>%
  filter(time %in% c("Start", "End")) %>%
  mutate(p_pos = p_m + p_s) %>%
  group_by(outcome) %>%
  summarize(
    p_pos = max(p_pos)
  ) %>%
  ungroup() %>%
  mutate(p_pos = p_pos + 5)

pval_ypos_post <- qol_imp_prop %>%
  filter(time %in% c("Post", "End")) %>%
  mutate(p_pos = p_m + p_s) %>%
  group_by(outcome) %>%
  summarize(
    p_pos = max(p_pos)
  ) %>%
  ungroup() %>%
  mutate(p_pos = p_pos + 5)

pval_ypos <- rbind(
  pval_ypos_end %>% mutate(time = "End"),
  pval_ypos_post %>% mutate(time = "Post")
) %>%
  arrange(outcome, time) %>%
  pull(p_pos)

# plot
pl_comp <- qol_imp_prop %>%
  ggplot(aes(x = outcome, fill = time, y = p_m)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = .66),
    width = .5
  ) +
  geom_errorbar(
    aes(ymin = p_m - p_s, ymax = p_m + p_s),
    position = position_dodge(width = .66),
    width = .2
  ) +
  ggsignif::geom_signif(
    annotations = qol_bin_pp,
    y_position = pval_ypos,
    xmin = c(.75, 1, 1.75, 2, 2.75, 3, 3.75, 4, 4.75, 5, 5.75, 6, 6.75, 7),
    xmax = c(1, 1.25, 2, 2.25, 3, 3.25, 4, 4.25, 5, 5.25, 6, 6.25, 7, 7.25),
    textsize = 8 / cm(1)
  ) +
  scale_y_continuous(
    breaks = scales::breaks_pretty(5),
    expand = expansion(mult = c(0, 0.025))
  ) +
  scale_fill_brewer(palette = "Set2") +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  labs(y = "Proportion (%)", x = "Pericardial changes") +
  theme_custom(8) +
  theme(
    legend.position = "bottom", legend.title = element_blank(),
    axis.title.x = element_blank(),
    plot.title = element_text(face = 1, size = 8, margin = margin(b = 20)),
    strip.text = element_text(face = 2),
    legend.key.height = unit(0.3, "cm"),
    legend.key.width = unit(0.5, "cm"),
    plot.margin = margin(l = 20, t = 5.5, b = 5.5, r = 5.5)
  )

save_plot(
  pl_comp,
  pdf_file = "results/qol-impairment-prevalence-by-time.pdf",
  w = 20, h = 10
)
```