---
title: "Analysis"
author: "Nicolas Banholzer"
date: "2023-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = "..")

re_fit <- FALSE
```

## Libraries

Loading packages...

```{r libraries, include=FALSE}
library(tidyverse)
library(reshape2)
library(rstanarm)
library(brms)
library(rstan)
library(tidybayes)
library(mice)
library(klaR)
source("utils/plotting.R")
source("utils/summaries.r")
```


## Data

Loading data...

```{r data, include=FALSE}
# data of patients with at least one outcome
qol <- readRDS("data-clean/phys-ment-data.rds") %>%
  arrange(time, record_id)
qol_start <- filter(qol, time == "Start treatment")
qol_end <- filter(qol, time == "End treatment")
qol_post <- filter(qol, time == "Post treatment")

# labeling
outcome_cont <- c(
  "phq9_score" = "PHQ9-S",
  "sf12_ment" = "QoL-MCS",
  "sf12_phys" = "QoL-PCS",
  "smwt_dist" = "6MWT-D",
  "stst_nr" = "STST-R"
)

outcome_bin <- c(
  "phq9_score_bin" = "PHQ9-S>10",
  "sf12_ment_bin" = "QoL-MCS<42",
  "sf12_phys_bin" = "QoL-PCS<50",
  "smwt_dist_bin" = "6MWT-D<400m",
  "stst_nr_bin" = "STST-R<20"
)

predictor <- c(
  "age" = "Age",
  "sex" = "Female",
  "hiv" = "HIV+",
  "mdr" = "MDR",
  "cavity" = "Cavitation"
)
predictor_2 <- predictor
predictor_2[1] <- "Age<30"

sites <- c("South Africa", "Malawi", "Mosambique", "Zambia", "Zimbabwe")

treat_effect_lab <- c("End vs Start treatment", "Post vs End treatment")
```

**Outcomes**:

The quality of life **(QoL)** outcomes:

* PHQ9-S: Patient Health Questionnaire 9 - Score
* QoL-MCS: Quality of Life - Mental Component Score
* QoL-PCS: Quality of Life - Physical Component Score
* 6MWT-D: 6 Minute Walk Test - Distance
* STST-R: Sit-to-Stand Test - Number of Repetitions

Respective we define impaired QoL as the following binary outcomes:

1. PHQ9-S > 10
2. QoL-MCS < 42
3. QoL-PCS < 50
4. 6MWT-D < 400m
5. STST-R < 20


**Exclusion:**

We exclude observations where we still wait for the next clinical visit.
That is, end of treatment or post treatment is still within 8mo since last visit.
Patients who died were only considered once at the the time of their death.
For example, if a patient died during treatment, we only consider the start and end of treatment, not post treatment.


**Missing observations**:

The outcomes can be missing for one of the following reasons:

1. *Death*: The patient died.
2. *LTFU*: The patient was lost to follow-up (LTFU).
3. *Unrecorded*: The outcome was not recorded.


**Patient characteristics**:

The predictors we consider are:

1. Age
2. sex
3. HIV test result
4. Multi-drug resistant TB (MDR-TB)
5. Presence of cavitation on chest X-ray.


**TB treatment outcome**:

The association with TB treatment outcome is currently not analyzed because it is mostly missing. 
Also, a treatment failure can result in death. 
Coding death as impaired QoL, dead patients would need to be excluded from the analysis.

## Descriptives

### Study population

```{r sample}
print(
  sprintf(
    "Total: %i patients with %i observations.",
    n_distinct(qol$record_id), nrow(qol)
  )
)
```

**Figure:** Number of observations by time.

```{r observations}
qol %>%
  group_by(time) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x = time, y = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), size = 8 / cm(1), vjust = -1) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(y = "No. of observations") +
  theme_custom() +
  theme(
    axis.title.x = element_blank()
  )
```

**Figure**: Proportion of patients lost to follow-up (LTFU).

```{r lost-to-follow-up}
qol %>%
  group_by(time) %>%
  summarize(
    `no. LTFU` = sum(ltfu),
    `proportion` = paste0(round(sum(ltfu) / n() * 100), "%")
  )
```

**Table**: Patient characteristics.

```{r predictors}
qol_start %>%
  dplyr::select(all_of(names(predictor))) %>%
  gather() %>%
  mutate(
    key = recode(key, !!!predictor),
    key = factor(key, levels = predictor)
  ) %>%
  group_by(key) %>%
  summarize_all(
    ~ ifelse(max(.x, na.rm = T) > 1, median_iqr(.x), count_bin(.x))
  ) %>%
  ungroup() %>%
  arrange(key) %>%
  set_names(c("Variable", "Summary"))
```

**Figure**: Share of sites.

```{r sites}
qol_start %>%
  group_by(site) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  ggplot(aes(x = site, y = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), size = 8 / cm(1), vjust = -1) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(y = "No. of patients") +
  theme_custom() +
  theme(
    axis.title.x = element_blank()
  )
```

**Table**: TB Treatment outcomes.

```{r tb-treatment-outcomes}
qol %>%
  filter(time == "End treatment") %>%
  summarize(`treatment success` = count_bin(treat_success))
```

### Outcomes

#### Summary

**Table**: Summary of continuous QoL outcomes.

```{r summaries}
# start of treatment
qol_start %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!!outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment start"))

# end of treatment
qol_end %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!!outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment end"))

# post treatment
qol_post %>%
  dplyr::select(all_of(names(outcome_cont))) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!!outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) post treatment"))
```

#### Distribution

**Figure**: Distribution of continuous QoL outcomes, excluding missing observations.

```{r density-distribution}
qol_denesity <- qol %>%
  dplyr::select(all_of(c("time", names(outcome_cont)))) %>%
  melt("time") %>%
  rename(outcome = variable) %>%
  mutate(outcome = recode(outcome, !!!outcome_cont)) %>%
  ggplot(aes(x = value, fill = time)) +
  facet_wrap(~outcome, scales = "free") +
  geom_density(alpha = .5) +
  coord_cartesian(expand = FALSE) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  labs(y = "Density") +
  theme_custom() +
  theme(
    legend.position = c(.85, .25),
    legend.direction = "vertical",
    legend.title = element_blank(),
    axis.title.x = element_blank()
  )

save_plot(
  qol_denesity,
  pdf_file = "results/outcome-density-distribution.png",
  w = 16, h = 12
)
```

**Figure:** Proportion of patients with impaired QoL.

```{r proportions}
var_names <- c(
  "Y" = "Yes",
  "N" = "No",
  "D" = "Death",
  "L" = "LTFU",
  "U" = "Unrecorded"
)

qol_bin_comp <- qol %>%
  dplyr::select(all_of(c("time", "ltfu", "death", names(outcome_bin)))) %>%
  melt(c("time", "ltfu", "death")) %>%
  group_by(time, variable) %>%
  summarize(
    Y = sum(value, na.rm = TRUE),
    N = sum(value == 0, na.rm = TRUE),
    D = sum(death),
    U = sum(is.na(value) & !ltfu & !death, na.rm = TRUE),
    L = sum(ltfu, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  rename(outcome = variable) %>%
  melt(c("time", "outcome")) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin),
    variable = recode(variable, !!!var_names),
    variable = factor(variable, levels = var_names)
  ) %>%
  group_by(time, outcome) %>%
  mutate(p = value / sum(value) * 100) %>%
  ungroup()

qol_bin_comp_pl <- ggplot(data = qol_bin_comp, mapping = aes(x = time, y = p)) +
  facet_wrap(~outcome, nrow = 1) +
  geom_bar(
    aes(fill = variable),
    stat = "identity",
    position = position_stack(reverse = TRUE)
  ) +
  geom_text(
    data = filter(qol_bin_comp, variable == "Yes"),
    mapping = aes(group = time, label = paste0(round(p), "%")),
    vjust = -.5, size = 8 / cm(1),
    color = wes_palette("Royal2")[1],
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = wes_palette("Royal2"),
  ) +
  scale_x_discrete(labels = c("Start", "End", "Post")) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  labs(
    y = "Proportion of patients (%)",
    x = "Treatment",
    fill = "a"
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_legend(title = element_blank())
  )

save_plot(
  qol_bin_comp_pl,
  pdf_file = "results/impaired-qol.png",
  w = 16, h = 10
)
```

#### Changes

**Figure**: Changes in QoL. Deaths imputed as impaired QoL. Unknown if QoL unrecorded or waiting for next clinical visit.

```{r changes}
qol_changes <- qol %>%
  dplyr::select(
    all_of(c("record_id", "time", "ltfu", "death", names(outcome_bin)))
  ) %>%
  mutate(across(
    all_of(names(outcome_bin)),
    ~ ifelse(death, 1, .x)
  )) %>%
  dplyr::select(-death, -ltfu) %>%
  melt(c("record_id", "time")) %>%
  dcast(record_id + variable ~ time) %>%
  mutate(
    change_end = `End treatment` - `Start treatment`,
    change_post = `Post treatment` - `End treatment`
  ) %>%
  rename(outcome = variable) %>%
  mutate(
    change_end =
      ifelse(is.na(change_end), "Unknown",
        ifelse(change_end == -1, "Improved",
          ifelse(change_end == 1, "Worsened",
            ifelse(`Start treatment` == 1, "No improvement", "Stable")
          )
        )
      ),
    change_post =
      ifelse(is.na(change_post), "Unknown",
        ifelse(change_post == -1, "Improved",
          ifelse(change_post == 1, "Worsened",
            ifelse(`End treatment` == 1, "No improvement", "Stable")
          )
        )
      )
  ) %>%
  dplyr::select(record_id, outcome, change_end, change_post) %>%
  melt(c("record_id", "outcome")) %>%
  mutate(
    value = factor(
      value,
      levels = c(
        "Worsened", "No improvement",
        "Unknown",
        "Stable", "Improved"
      )
    ),
    outcome = recode(gsub("_bin", "", outcome), !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  )

qol_changes_sum <- qol_changes %>%
  group_by(outcome, variable, value) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(outcome, variable) %>%
  mutate(p = 100 * n / sum(n)) %>%
  ungroup() %>%
  mutate(variable = ifelse(
    variable == "change_end",
    treat_effect_lab[1],
    treat_effect_lab[2]
  ), variable = factor(variable, levels = treat_effect_lab))

change_cols <- RColorBrewer::brewer.pal(5, "RdYlBu")
change_cols[3] <- "grey"

qol_changes_sum_pl <- ggplot(
  data = qol_changes_sum,
  mapping = aes(x = outcome, y = p, fill = value)
) +
  facet_wrap(~variable, nrow = 1) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    data = filter(qol_changes_sum, value == "Improved"),
    mapping = aes(group = value, label = paste0(round(p), "%")),
    vjust = -.5, size = 8 / cm(1),
    color = change_cols[5],
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = change_cols,
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  labs(
    y = "Proportion of patients (%)",
    x = "Treatment", fill = "a", alpha = "a"
  ) +
  theme_custom() +
  theme(legend.position = "bottom") +
  guides(
    fill = guide_legend(title = element_blank()),
    alpha = guide_legend(title = element_blank())
  )

save_plot(
  qol_changes_sum_pl,
  pdf_file = "results/changes-in-qol.png",
  w = 16, h = 10
)
```

#### Stratum flow 

**Figure**: Patient flow between stratums.

```{r flow, fig.height=10}
# define categories
sankey_cats <- data.frame(
  outcome = outcome_cont,
  cutoff = c(10, 42, 50, 400, 20)
) %>%
  mutate(
    yes = paste(outcome, "smaller", cutoff),
    no = paste(outcome, "bigger", cutoff),
    death = paste(outcome, "death"),
    ltfu = paste(outcome, "ltfu"),
    unknown = paste(outcome, "unknown")
  ) %>%
  dplyr::select(yes, no, death, ltfu, unknown) %>%
  gather() %>%
  mutate(
    names = gsub("smaller \\d*", "yes", value),
    names = gsub("bigger \\d*", "no", names),
    value = gsub("smaller", "<", value),
    value = gsub("bigger", ">", value)
  )
sankey_cats_vec <- sankey_cats$value
names(sankey_cats_vec) <- sankey_cats$names

# create links
links <- qol %>%
  dplyr::select(
    all_of(c("time", "record_id", "death", "ltfu", names(outcome_bin)))
  ) %>%
  reshape2::melt(c("time", "record_id", "death", "ltfu")) %>%
  rename(outcome = variable) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin),
    value =
      ifelse(death == 1 & is.na(value) & time != "Start treatment", "Death",
        ifelse(ltfu, "LTFU",
          ifelse(is.na(value), "Unrecorded",
            ifelse(value == 1, "Yes", "No")
          )
        )
      ),
    value = factor(value, levels = rev(var_names))
  ) %>%
  group_by(outcome, time, value) %>%
  mutate(freq = n()) %>%
  ungroup()

flow_pl <- links %>%
  ggplot(aes(
    x = time,
    stratum = value,
    alluvium = record_id,
    fill = value,
    label = freq
  )) +
  facet_wrap(~outcome) +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft") +
  geom_stratum() +
  geom_text(stat = "stratum", size = 8 / cm(1)) +
  scale_fill_manual(values = rev(wes_palette("Royal2"))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.01))) +
  scale_x_discrete(labels = c("Start", "End", "Post")) +
  labs(x = "Treatment") +
  theme_custom() +
  theme(
    legend.title = element_blank(),
    legend.position = c(.8, .2),
    axis.title.y = element_blank()
  )

save_plot(
  flow_pl,
  pdf_file = "results/sankey.png",
  w = 16, h = 20
)
```

## Imputation

Before modeling, we impute the missing values in the binary QoL outcome.
We will exclude LTFU, thereby assuming they are similar to the observed patients.
Deaths will be imputed as having the health condition. 

Thus, we will subsequently impute the "unrecorded" outcomes. 
For this, we use mice (Multivariate Imputation by Chained Equations).
Imputation will be based on the other outcomes, treatment time, and patient characteristics.

**Figure**: Imputation check for a random sample of 40 patients, showing the imputed majority label across datasets.

```{r imputation, fig.height=12}
# wide format
qol_excl_ltfu <- qol %>%
  # impute deaths
  mutate(across(
    all_of(names(outcome_bin)),
    ~ ifelse(death, 1, .x)
  )) %>%
  # exclude ltfu
  filter(!ltfu) %>%
  ungroup() %>%
  mutate(
    age = ifelse(age < 30, 1, 0),
    across(c(-record_id, -time), ~ ifelse(.x == 1, "Yes", "No")),
    across(c(-record_id, -time), ~ factor(.x, levels = c("No", "Yes")))
  ) %>%
  dplyr::select(all_of(c(
    "record_id",
    "time",
    names(outcome_bin),
    names(predictor)
  )))

# imputation
n_datasets <- 10
qol_excl_ltfu_imp <- mice(
  qol_excl_ltfu[, -1],
  m = n_datasets,
  maxit = 10,
  seed = 12345,
  print = FALSE
)
# plot(qol_excl_ltfu_imp, layout = c(2, 8))

# data
qol_imp <- complete(qol_excl_ltfu_imp, "long") %>%
  mutate(record_id = rep(qol_excl_ltfu$record_id, times = n_datasets))

# check majority label
stat_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

qol_excl_ltfu_imp_mode <- qol_imp %>%
  dplyr::select(all_of(c(".imp", "record_id", "time", names(outcome_bin)))) %>%
  melt(c(".imp", "record_id", "time")) %>%
  group_by(record_id, time, variable) %>%
  summarize(
    value = stat_mode(value)
  ) %>%
  rename(imputed = value) %>%
  left_join(
    qol_excl_ltfu %>%
      dplyr::select(all_of(c("record_id", "time", names(outcome_bin)))) %>%
      melt(c("record_id", "time")) %>%
      rename(observed = value),
    by = c("record_id", "time", "variable")
  ) %>%
  mutate(is_observed = ifelse(is.na(observed), FALSE, TRUE)) %>%
  rename(outcome = variable)

check_imp_pl <- qol_excl_ltfu_imp_mode %>%
  group_by(record_id) %>%
  filter(any(!is_observed)) %>%
  ungroup() %>%
  filter(record_id %in% sample(unique(record_id), 40)) %>%
  ggplot(aes(y = time, x = outcome, color = is_observed, label = imputed)) +
  geom_text(size = 8 / cm(1)) +
  facet_wrap(~record_id, ncol = 4) +
  labs(color = "Is observed") +
  scale_x_discrete(labels = function(x) outcome_bin) +
  scale_y_discrete(labels = function(x) c("Start", "End", "Post")) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "top",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

save_plot(
  check_imp_pl,
  pdf_file = "results/imputation-check.png",
  w = 16, h = 30
)
```


## Modeling

### Association with patient characteristics at baseline

The association between patient characteristics and QoL at the start of treatment is estimated using a logistic regression model.
We use a multilevel model, which partially pools the association across outcomes. 
Furthermore, the estimation is performed on each imputed dataset and the estimates are pooled across datasets. 

**Figure**: Estimated association of QoL at baseline with patient characteristics.

```{r association-baseline}
# prepare data
qol_start_long <- qol_imp %>%
  filter(time == "Start treatment") %>%
  dplyr::select(all_of(c(
    ".imp",
    "record_id",
    names(outcome_bin),
    names(predictor)
  ))) %>%
  mutate(across(
    c(-`.imp`, -`record_id`),
    ~ ifelse(.x == "Yes", 1, 0)
  )) %>%
  melt(c("record_id", ".imp", names(predictor))) %>%
  nest(data = -c(`.imp`))

# estimate models
if (re_fit) {
  assoc_base_mod <- brm_multiple(
    data = qol_start_long$data,
    formula = value ~ age + sex + hiv + mdr + cavity +
      (1 + age + sex + hiv + mdr + cavity | variable),
    family = bernoulli(link = "logit"),
    prior = c(
      prior(normal(0, 1), class = sd),
      prior(normal(0, 2.5), class = b),
      prior(normal(0, 10), class = Intercept)
    ),
    chains = 4, iter = 2000,
    seed = 12345,
    cores = 4
  )

  saveRDS(assoc_base_mod, "fitted-models/assoc_base_mod.rds")
} else {
  assoc_base_mod <- readRDS("fitted-models/assoc_base_mod.rds")
}


# posterior draws
assoc_base_draws_pop <- assoc_base_mod %>%
  spread_draws(
    b_age,
    b_sex,
    b_hiv,
    b_mdr,
    b_cavity
  ) %>%
  melt(c(".chain", ".iteration", ".draw")) %>%
  rename(predictor = variable) %>%
  mutate(predictor = gsub("b_", "", predictor))

assoc_base_draws_re <- assoc_base_mod %>%
  spread_draws(
    r_variable[outcome, predictor]
  ) %>%
  rename(value = r_variable)

assoc_base_draws <- left_join(
  assoc_base_draws_pop,
  assoc_base_draws_re,
  by = c(".chain", ".iteration", ".draw", "predictor")
) %>%
  mutate(
    value = value.x + value.y,
    value = exp(value),
    predictor = recode(predictor, !!!predictor_2),
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin),
    predictor = factor(predictor, levels = predictor_2)
  )

# plot associations
assoc_base_pl <- assoc_base_draws %>%
  ggplot(aes(x = value, y = outcome)) +
  facet_wrap(~predictor, ncol = 2) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_pointinterval(
    .width = c(.5, 0.95),
    point_size = 1,
    shape = 21, point_fill = "white"
  ) +
  labs(
    x = "Adj. Odds Ratio",
  ) +
  theme_custom() +
  theme(axis.title.y = element_blank())

save_plot(
  assoc_base_pl,
  pdf_file = "results/association-predictors-baseline.png",
  w = 16, h = 10
)
```

### QoL impairment over time

The proportion of patients with impaired QoL for each outcome is estimated using binomial models.
Again, the estimates are pooled across the imputed datasets.
Patients lost-to-follow-up are excluded from the analysis, i.e. assuming they are similar to the studied population.

**Figure**: Estimated treatment effects. 

```{r treatment-effects}
# create binomial datasets
qol_binom <- qol_imp %>%
  dplyr::select(all_of(c(".imp", "record_id", "time", names(outcome_bin)))) %>%
  melt(c(".imp", "record_id", "time")) %>%
  rename(outcome = variable) %>%
  group_by(`.imp`, time, outcome) %>%
  summarize(
    n = n(),
    x = sum(value == "Yes")
  ) %>%
  ungroup() %>%
  mutate(p = x / n) %>%
  nest(data = c(-`.imp`))

# estimate models
if (re_fit) {
  binom_mod <- brm_multiple(
    data = qol_binom$data,
    formula = bf(x | trials(n) ~ 1 + time + (1 + time | outcome)),
    family = binomial(link = "logit"),
    prior = c(
      prior(normal(0, 1), class = sd),
      prior(normal(0, 10), class = Intercept)
    ),
    chains = 4,
    iter = 2000,
    seed = 12345,
    cores = 4
  )

  saveRDS(binom_mod, "fitted-models/mod_binom.rds")
} else {
  binom_mod <- readRDS("fitted-models/mod_binom.rds")
}

# treatment effects
treat_effect_draws_pop <- binom_mod %>%
  spread_draws(
    b_timeEndtreatment,
    b_timePosttreatment
  ) %>%
  melt(c(".chain", ".iteration", ".draw")) %>%
  rename(time = variable) %>%
  mutate(time = gsub("b_", "", time))

treat_effect_draws_re <- binom_mod %>%
  spread_draws(
    r_outcome[outcome, time]
  ) %>%
  rename(value = r_outcome) %>%
  filter(time != "Intercept")

treat_effect_draws <- left_join(
  treat_effect_draws_pop, treat_effect_draws_re,
  by = c(".chain", ".iteration", ".draw", "time")
) %>%
  mutate(value = value.x + value.y) %>%
  dplyr::select(-`value.x`, -`value.y`) %>%
  dcast(`.chain` + `.iteration` + `.draw` + outcome ~ time) %>%
  mutate(timePosttreatment = timePosttreatment - timeEndtreatment) %>%
  melt(c(".chain", ".iteration", ".draw", "outcome")) %>%
  rename(time = variable) %>%
  mutate(
    value = exp(value),
    time = ifelse(
      time == "timeEndtreatment",
      treat_effect_lab[1],
      treat_effect_lab[2]
    ),
    time = factor(time, levels = treat_effect_lab),
    outcome = recode(outcome, !!!outcome_bin),
    outcome = factor(outcome, levels = outcome_bin)
  )

treat_effects_pl <- treat_effect_draws %>%
  ggplot(aes(x = value, y = outcome)) +
  facet_wrap(~time) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  scale_x_continuous(limits = c(0, NA)) +
  labs(x = "Estimated Adj. OR") +
  stat_pointinterval(point_size = 2) +
  theme_custom() +
  theme(axis.title.y = element_blank())

save_plot(
  treat_effects_pl,
  pdf_file = "results/treatment-effects.png",
  w = 16, h = 6
)
```

**Figure**: Estimated proportion of patients with impaired QoL. Pr = Bayesian probability that the proportions are unequal.

```{r proportions-impaired-qol}
# binary changes
qol_binom_pred <- qol_binom$data[[1]] %>%
  modelr::data_grid(time, outcome, n = 100) %>%
  add_epred_draws(binom_mod) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_bin)
  ) %>%
  ungroup() %>%
  mutate(p = `.epred` / n) %>%
  dplyr::select(-`.chain`, -`.iteration`, -n, -`.row`, -`.epred`) %>%
  rename(draw = `.draw`)

# comparison
qol_pred_comp <- qol_binom_pred %>%
  dcast(outcome + draw ~ time) %>%
  mutate(
    es1 = `End treatment` < `Start treatment`,
    es2 = `End treatment` > `Start treatment`,
    ep1 = `Post treatment` < `End treatment`,
    ep2 = `Post treatment` > `End treatment`
  ) %>%
  group_by(outcome) %>%
  summarise(across(c(es1, es2, ep1, ep2), ~ sum(.x) / n())) %>%
  rowwise() %>%
  mutate(
    es = pmax(es1, es2),
    ep = pmax(ep1, ep2)
  ) %>%
  ungroup() %>%
  dplyr::select(outcome, es, ep) %>%
  melt("outcome") %>%
  rename(comparison = variable) %>%
  mutate(comparison = as.character(comparison))

# annotation
qol_pred_comp_annon <- data.frame(
  outcome = outcome_bin,
  xmin_es = c(0.75, 1.75, 2.75, 3.75, 4.75),
  xmin_ep = 1:5
) %>%
  mutate(
    xmax_es = xmin_es + 0.25,
    xmax_ep = xmin_ep + 0.25
  ) %>%
  mutate(
    ypos_es = qol_binom_pred %>%
      filter(time %in% c("Start treatment", "End treatment")) %>%
      group_by(outcome, time) %>%
      summarize(ypos = quantile(p, .975)) %>%
      ungroup() %>%
      group_by(outcome) %>%
      summarize(ypos = max(ypos)) %>%
      ungroup() %>%
      dplyr::select(ypos) %>%
      unlist(),
    ypos_ep = qol_binom_pred %>%
      filter(time %in% c("End treatment", "Post treatment")) %>%
      group_by(outcome, time) %>%
      summarize(ypos = quantile(p, .975)) %>%
      ungroup() %>%
      group_by(outcome) %>%
      summarize(ypos = max(ypos)) %>%
      ungroup() %>%
      dplyr::select(ypos) %>%
      unlist()
  ) %>%
  melt("outcome") %>%
  mutate(
    comparison = stringi::stri_extract(variable, regex = "_.*"),
    variable = gsub("_.*", "", variable),
    comparison = gsub("_", "", comparison)
  ) %>%
  dcast(outcome + comparison ~ variable)

qol_pred_comp <- qol_pred_comp %>%
  left_join(qol_pred_comp_annon, by = c("outcome", "comparison")) %>%
  mutate(
    lab = round(value * 100),
    lab = paste0("Pr.", ifelse(lab == 100, ">99", paste0("=", lab)), "%")
  ) %>%
  mutate(ypos = ypos + .05)

# plot
qol_pred_change_pl <- qol_binom_pred %>%
  ggplot(aes(x = outcome, y = p, color = time)) +
  stat_interval(aes(fill_ramp = after_stat(level)),
    position = position_dodge(width = .75),
    .width = c(.5, .95),
    alpha = .5
  ) +
  stat_pointinterval(
    .width = NA,
    position = position_dodge(width = .75),
    shape = 21, fill = "white", size = 2,
    show.legend = FALSE
  ) +
  ggsignif::geom_signif(
    annotations = qol_pred_comp$lab,
    y_position = qol_pred_comp$ypos,
    xmin = qol_pred_comp$xmin,
    xmax = qol_pred_comp$xmax,
    textsize = 8 / cm(1), color = "black"
  ) +
  scale_color_manual(values = wes_palette("Moonrise2")) +
  scale_y_continuous(
    expand = expansion(add = c(0, 0.05)),
    limits = c(0, 1.05),
    breaks = seq(0, 1, .25),
    labels = function(x) x * 100
  ) +
  labs(
    y = "Proportion of patients (%)"
  ) +
  theme_custom() +
  theme(
    legend.position = "bottom",
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.key.width = unit(.3, "cm"), legend.key.height = unit(.6, "cm")
  )

save_plot(
  qol_pred_change_pl,
  pdf_file = "results/treatment-binomial-comparison.png",
  w = 16, h = 10
)
```

### Clustering treatment changes

Since there only few changes post treatment, we focus subsequently on the changes between start and end of treatment.
For that purpose, we cluster the patients based on the changes in health conditions.
Again, we exclude patients lost-to-follow-up.

**Figure**: Changes in QoL through treatment for random sample of 40 patients.

```{r individual-heatmap, fig.height=10}
qol_treat_changes <- qol %>%
  dplyr::filter(time %in% c("Start treatment", "End treatment")) %>%
  mutate(across(
    all_of(names(outcome_bin)),
    ~ ifelse(death, 1, .x)
  )) %>%
  group_by(record_id) %>%
  filter(
    n() == 2,
    !any(ltfu)
  ) %>%
  ungroup() %>%
  dplyr::select(record_id, time, all_of(names(outcome_bin))) %>%
  melt(c("record_id", "time")) %>%
  dcast(record_id + variable ~ time) %>%
  mutate(
    change = `End treatment` - `Start treatment`
  ) %>%
  rename(outcome = variable) %>%
  mutate(
    value =
      ifelse(is.na(change), NA,
        ifelse(change == -1, "Improved",
          ifelse(change == 1, "Worsened",
            ifelse(`Start treatment` == 1, "No improvement", "Stable")
          )
        )
      ),
    value = factor(value, levels = c("Worsened", "No improvement", "Stable", "Improved")), # nolint
    outcome = recode(gsub("_bin", "", outcome), !!!outcome_cont),
    outcome = factor(outcome, levels = outcome_cont)
  )

qol_treat_changes_id_pl <- qol_treat_changes %>%
  filter(record_id %in% sample(record_id, 40)) %>%
  ggplot(aes(x = outcome, y = record_id, fill = value)) +
  geom_tile() +
  scale_fill_brewer(palette = "RdYlBu", na.value = "grey") +
  coord_cartesian(expand = FALSE) +
  theme_custom() +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "top"
  )

save_plot(
  qol_treat_changes_id_pl,
  pdf_file = "results/treat-change-individual-heatmap.png",
  w = 16, h = 25
)
```

We use the k-modes clustering algorithm, which is suitable for categorical data.
The clustering is performed on the imputed data.
We use the majority label for each imputed observation across datasets.
The reason is that cluster labels are assigned manually, and it is easier to do this once than for all datasets separately.

We tested various numbers of clusters.
We could most clearly distinguish between two broad clusters.
One cluster showed a majority of patients with stable or improved health conditions.
The other cluster showed a majority of patients with stagnation or deterioration in health conditions.

**Figure**: Clusters of changes in QoL through treatment.

```{r clustering}
# clustering
treat_change_cat <- qol_excl_ltfu_imp_mode %>%
  filter(time %in% c("Start treatment", "End treatment")) %>%
  mutate(imputed = ifelse(imputed == "Yes", 1, 0)) %>%
  dplyr::select(record_id, time, outcome, imputed) %>%
  group_by(record_id, outcome) %>%
  filter(n() == 2) %>%
  ungroup() %>%
  dcast(record_id + outcome ~ time) %>%
  mutate(
    change = `End treatment` - `Start treatment`
  ) %>%
  mutate(
    value =
      ifelse(change == -1, "Improved",
        ifelse(change == 1, "Worsened",
          ifelse(`Start treatment` == 1, "No improvement", "Stable")
        )
      ),
    value = factor(value, levels = c("Worsened", "No improvement", "Stable", "Improved")), # nolint
    value_int = as.integer(value)
  ) %>%
  dplyr::select(record_id, outcome, value_int) %>%
  dcast(record_id ~ outcome)

treat_change_cl <- klaR::kmodes(
  treat_change_cat[, -1],
  modes = 2,
  iter.max = 100,
  weighted = FALSE,
  fast = TRUE
)

# plot cluster
treat_change_cat_cl <- cbind(
  treat_change_cat,
  cluster = treat_change_cl$cluster
) %>%
  melt(c("record_id", "cluster")) %>%
  mutate(value = recode(value, !!!c(
    "1" = "Worsened",
    "2" = "No improvement",
    "3" = "Stable",
    "4" = "Improved"
  )), value = factor(
    value,
    levels = c(
      "Worsened",
      "No improvement",
      "Stable",
      "Improved"
    )
  ))

treat_change_cat_cl_sum <- treat_change_cat_cl %>%
  rename(outcome = variable) %>%
  group_by(cluster, outcome, value) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(cluster, outcome) %>%
  mutate(p = 100 * n / sum(n)) %>%
  ungroup()

treat_change_cl_pl <- treat_change_cat_cl_sum %>%
  ggplot(aes(x = outcome, y = p, fill = value)) +
  facet_wrap(~ paste("Cluster", cluster), ncol = 2) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    y = "Proportion of patients (%)",
    fill = "Health condition"
  ) +
  scale_fill_brewer(palette = "RdYlBu") +
  scale_x_discrete(labels = function(x) outcome_bin) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_custom() +
  theme(
    legend.position = "top",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

save_plot(
  treat_change_cl_pl,
  pdf_file = "results/treatment-change-cluster.png",
  w = 16, h = 8
)

# label cluster
tc_cl_name <- treat_change_cat_cl_sum %>%
  group_by(cluster) %>%
  summarize(
    Improved = sum(p[value == "Stable"]),
    Worsened = sum(p[value == "No improvement"]),
  ) %>%
  ungroup() %>%
  melt("cluster") %>%
  group_by(variable) %>%
  mutate(r = rank(value)) %>%
  ungroup() %>%
  filter(r == 2)
tc_cl_name_vec <- tc_cl_name$variable
names(tc_cl_name_vec) <- tc_cl_name$cluster
```

**Figure**: Estimated association of patient characteristics with stagnation (no improvement) in QoL.

```{r association-patient-characteristics}
# add baseline characteristics
treat_change_cl_id <- treat_change_cat_cl %>%
  mutate(cluster = recode(cluster, !!!tc_cl_name_vec)) %>%
  mutate(stagnation = ifelse(cluster == "Worsened", 1, 0)) %>%
  group_by(record_id) %>%
  slice(1) %>%
  ungroup() %>%
  left_join(
    qol_start %>%
      dplyr::select(all_of(c("record_id", names(predictor)))),
    by = "record_id"
  ) %>%
  mutate(age = ifelse(age <= 30, 1, 0)) %>%
  ungroup()

# estimate association
tc_assoc <- rstanarm::stan_glm(
  stagnation ~ age + sex + hiv + mdr + cavity,
  data = treat_change_cl_id,
  family = binomial(link = "logit"),
  seed = 12345,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores = 4
)

# plot association
tc_assoc_pl <- spread_draws(
  tc_assoc,
  age, sex, hiv, mdr, cavity
) %>%
  dplyr::select(age, sex, hiv, mdr, cavity) %>%
  gather() %>%
  mutate(
    key = recode(key, !!!predictor_2),
    key = factor(key, levels = predictor_2),
    value = exp(value)
  ) %>%
  ggplot(aes(x = value, y = key)) +
  geom_vline(aes(xintercept = 1), linetype = "dotted", color = "red") +
  stat_pointinterval(
    point_size = 2, .width = c(.5, .95),
    shape = 21, point_fill = "white"
  ) +
  labs(
    x = "Adj. Odds Ratio"
  ) +
  theme_custom() +
  theme(axis.title.y = element_blank())

save_plot(
  tc_assoc_pl,
  pdf_file = "results/treatment-change-association.png",
  w = 8, h = 6
)
```

*Next step:* Evaluate the association of TB treatment success with stagnation in QoL.
However, treatment success is only available for a small subset of patients.
The subset becomes even smaller when excluding deaths, i.e. focusing on survivors.  
Thus, this analysis is currently not performed.

```{r association-treat-outcome}
# # subset of patients for which we have the treatment outcome
# tb_treat_out <- qol %>%
#   filter(
#     time == "End treatment",
#     !death,
#     !is.na(treat_success)
#   ) %>%
#   dplyr::select(record_id, treat_success) %>%
#   left_join(
#     treat_change_cl_id %>%
#       dplyr::select(record_id, stagnation),
#     by = "record_id"
#   )

# # estimate association
# tb_treat_out_assoc <- rstanarm::stan_glm(
#   improvement ~ treat_success,
#   data = tb_treat_out,
#   family = binomial(link = "logit"),
#   seed = 12345,
#   chains = 4,
#   iter = 2000,
#   warmup = 1000,
#   cores = 4
# )

# # report association
# b_treat_success <- spread_draws(
#   tb_treat_out_assoc,
#   treat_success
# ) %>%
#   mutate(treat_success = exp(treat_success)) %>%
#   median_qi()
```
