---
title: "Analysis"
author: "Nicolas Banholzer"
date: "2023-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Loading packages...

```{r, include=FALSE}
library(tidyverse)
library(reshape2)
library(rstanarm)
library(rstan)
library(tidybayes)
source("utils/plotting.R")
source("utils/summaries.r")
```


## Data

Loading data...

```{r, include=FALSE}
# data of patients with at least one outcome 
QoL <- readRDS("../data-clean/phys-ment-data.rds")
QoL_start <- filter(QoL, time == "Start treatment")
QoL_end <- filter(QoL, time == "End treatment")
QoL_post <- filter(QoL, time == "Post treatment")

# labeling
outcome_cont <- c("PHQ9-S", "QoL-MCS", "QoL-PCS", "6MWT-D", "STST-R")
names(outcome_cont) <- c("phq9_score", "sf12_ment", "sf12_phys", "smwt_dist", "stst_nr")
outcome_bin <- c("PHQ9-S<10", "QoL-MCS<42", "QoL-PCS<50", "6MWT-D<400m", "STST-R<20")
names(outcome_bin) <- c("phq9_score_bin", "sf12_ment_bin", "sf12_phys_bin", "smwt_dist_bin", "stst_nr_bin")
predictor <- c("Age", "Female", "Clinically diagnosed", "HIV+", "High bacterial load", "MDR", "Cavitation", "Opacity")
names(predictor) <- c("age", "sex", "clindiag", "hiv_test_result", "highbact", "drug_resistant", "cavity", "opacity")
```

* We analyze all patients with complete data, i.e. data recorded at treatment start & end and post treatment.
* Thereby, we also include patients that are lost to follow-up, i.e. no recorded data in 8 months since last visit.
* Outcomes for those lost to follow-up are treated as NA (not available).
* There may be further NAs for single outcomes, i.e. some endpoints not recorded / measured.

## Descriptives

### Study population

```{r, echo=F, warning=FALSE, message=FALSE}
print(sprintf("Total: %i patients with %i observations.", nrow(QoL) / 3, nrow(QoL)))
```

```{r, echo=F, warning=FALSE, message=FALSE}
# no ltfu
QoL %>%
  group_by(time) %>%
  summarize(p = sum(ltfu, na.rm = T) / n()) %>%
  ungroup() %>%
  ggplot(aes(x = time, y = p)) +
  geom_bar(stat = "identity") +
  scale_y_continuous(expand = c(0,0)) +
  labs(y = "Proportion of patients (%)",
       title = "Patients lost to follow-up (LTFU) over time",
       subtitle = "LTFU if >8months since last visit and no outcomes recorded") +
  theme_custom() +
  theme(axis.title.x = element_blank())
```

```{r, echo=F, warning=FALSE, message=FALSE}
# patient characteristics
QoL_start %>%
  dplyr::select(age, sex, hiv_test_result, drug_resistant, clindiag, cavity, opacity, highbact) %>%
  gather() %>%
  mutate(key = recode(key, !!! predictor),
         key = factor(key, levels = predictor)) %>%
  group_by(key) %>%
  summarize_all(~ ifelse(max(.x, na.rm = T) > 1, median_iqr(.x), count_per.bin(.x))) %>%
  ungroup() %>%
  arrange(key) %>%
  set_names(c("Variable", "Median (IQR) or N (%)"))
```

```{r, echo=F, warning=FALSE, message=FALSE}
# sites
count_per.cat(QoL_start$site)
```

### Outcomes

#### Summary

```{r, echo = F, warning=FALSE, message=FALSE}
# start of treatment
QoL_start %>%
  dplyr::select(sf12_phys, sf12_ment, phq9_score, smwt_dist, stst_nr) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!! outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment start"))

# end of treatment
QoL_end %>%
  dplyr::select(sf12_phys, sf12_ment, phq9_score, smwt_dist, stst_nr) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!! outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) at treatment end"))
  
# post treatment
QoL_post %>%
  dplyr::select(sf12_phys, sf12_ment, phq9_score, smwt_dist, stst_nr) %>%
  gather() %>%
  group_by(key) %>%
  summarize(median_and_iqr = median_iqr(value)) %>%
  ungroup() %>%
  mutate(key = recode(key, !!! outcome_cont)) %>%
  arrange(key) %>%
  set_names(c("Outcome", "Median (IQR) post treatment"))
```

#### Treatment comparison

```{r, echo = F, warning=FALSE, message=FALSE}
# comparison continuous data excluding NAs
QoL %>%
  dplyr::select(time, sf12_phys, sf12_ment, smwt_dist, stst_nr, phq9_score) %>%
  melt("time") %>%
  rename(outcome = variable) %>%
  mutate(outcome = recode(outcome, !!! outcome_cont)) %>%
  ggplot(aes(x = value, fill = time)) +
  facet_wrap(~ outcome, scales = "free") +
  geom_density(alpha = .5) +
  coord_cartesian(expand = F) +
  scale_fill_manual(values = wes_palette("Moonrise2")) +
  labs(title = "Physical and mental health scores through treatment",
       subtitle = "Health score distribution at treatment start & end and post treatment",
       y = "Density") +
  theme_custom() +
  theme(legend.position = c(.85, .25),
        legend.direction = "vertical",
        legend.title = element_blank(),
        axis.title.x = element_blank())
```

```{r, echo = F, warning=FALSE, message=FALSE}
# comparison binary data
QoL_bin_comp <- QoL %>%
  dplyr::select(time, ltfu, sf12_phys_bin, sf12_ment_bin, smwt_dist_bin, stst_nr_bin, phq9_score_bin) %>%
  melt(c("time", "ltfu"))  %>%
  group_by(time, variable) %>%
  summarize(Yes = sum(value, na.rm = T),
            No = sum(value == 0, na.rm = T),
            Unrecorded = sum(is.na(value) & !ltfu, na.rm = T),
            LTFU = sum(ltfu, na.rm = T)) %>%
  ungroup() %>%
  rename(outcome = variable) %>%
  melt(c("time", "outcome")) %>%
  mutate(outcome = recode(outcome, !!! outcome_bin),
         outcome = factor(outcome, levels = outcome_bin),
         variable = factor(variable, levels = c("LTFU", "Unrecorded", "No", "Yes"))) %>%
  group_by(time, outcome) %>%
  mutate(p = value / sum(value) * 100) %>%
  ungroup()

QoL_bin_comp_pl <- ggplot(data = QoL_bin_comp, mapping = aes(x = time, y = p)) +
  facet_wrap(~ outcome, nrow = 1) +
  geom_bar(mapping = aes(alpha = variable, fill = variable), stat = "identity") +
  geom_text(data = filter(QoL_bin_comp, variable == "Yes"), 
            mapping = aes(group = time, label = paste0(round(p), "%")), 
            vjust = -.5, size = 8 / cm(1), color = wes_palette("AsteroidCity2")[1], show.legend = F) +
  scale_fill_manual(values = wes_palette("AsteroidCity2"), breaks = c("Yes", "No", "Unrecorded", "LTFU")) +
  scale_x_discrete(labels = c("Start", "End", "Post")) +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  scale_alpha_discrete(range = c(.4, 1), breaks = c("Yes", "No", "Unrecorded", "LTFU")) +
  labs(y = "Proportion of patients (%)", x = "Treatment", fill = "a", alpha = "a",
       title = "Physical and mental health conditions through treatment",
       subtitle = "Proportion of patients with health conditions at treatment start & end and post treatment") +
  theme_custom() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(title = element_blank()),
         alpha = guide_legend(title = element_blank()))

save_plot(QoL_bin_comp_pl, pdf_file = "../results/score-treatment-comparison.pdf", w = 16, h = 10)
```

#### Changes

```{r, echo = F, warning=FALSE, message=FALSE}
QoL_changes <- QoL %>%
  dplyr::select(record_id, time, sf12_phys_bin, sf12_ment_bin, smwt_dist_bin, stst_nr_bin, phq9_score_bin) %>%
  melt(c("record_id", "time")) %>%
  dcast(record_id + variable ~ time) %>%
  mutate(change_end = `End treatment` - `Start treatment`,
         change_post = `Post treatment` - `End treatment`) %>%
  dplyr::select(record_id, variable, change_end, change_post) %>%
  rename(outcome = variable) %>%
  melt(c("record_id", "outcome")) %>%
  mutate(variable = ifelse(variable == "change_end", "End vs Start", "Post vs End"),
         value = ifelse(is.na(value), "Unknown", ifelse(value == -1, "Improved", ifelse(value == 0, "Unchanged", "Worsened"))),
         value = factor(value, levels = c("Unknown", "Worsened", "Unchanged", "Improved")),
         outcome = recode(gsub("_bin", "", outcome), !!! outcome_cont),
         outcome = factor(outcome, levels = outcome_cont)) 

QoL_changes_sum <- QoL_changes %>%
  group_by(outcome, variable, value) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(outcome, variable) %>%
  mutate(p = 100 * n / sum(n)) %>%
  ungroup()

QoL_changes_sum_pl <- ggplot(data = QoL_changes_sum, mapping = aes(x = outcome, y = p, fill = value)) +
  facet_wrap(~ variable, nrow = 1) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = filter(QoL_changes_sum, value == "Improved"), 
            mapping = aes(group = value, label = paste0(round(p), "%")), 
            vjust = -.5, size = 8 / cm(1), color = wes_palette("AsteroidCity2")[3], show.legend = F) +
  scale_fill_manual(values = c(rev(wes_palette("AsteroidCity2")[1:3]), "grey") , breaks = c("Improved", "Unchanged", "Worsened", "Unknown")) +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(0, NA), clip = "off") +
  labs(y = "Proportion of patients (%)", x = "Treatment", fill = "a", alpha = "a",
       title = "Changes in physical and mental health in TB patients",
       subtitle = "Proportion of patients by changes between treatment end vs start and post treatment vs treatment end") +
  theme_custom() +
  theme(legend.position = "bottom") +
  guides(fill = guide_legend(title = element_blank()),
         alpha = guide_legend(title = element_blank()))

QoL_changes_sum_pl
```

#### Sankey 

```{r, echo = F, warning=FALSE, message=FALSE}
sankey_cats <- data.frame(outcome = outcome_cont,
                          cutoff = c(10, 42, 50, 400, 20)) %>%
  mutate(yes = paste(outcome, "smaller", cutoff),
         no = paste(outcome, "bigger", cutoff),
         ltfu = paste(outcome, "ltfu"),
         unknown = paste(outcome, "unknown")) %>%
  dplyr::select(yes, no, ltfu, unknown) %>%
  gather() %>%
  mutate(names = gsub("smaller \\d*", "yes", value),
         names = gsub("bigger \\d*", "no", names),
         value = gsub("smaller", "<", value),
         value = gsub("bigger", ">", value)) 
sankey_cats_vec <- sankey_cats$value
names(sankey_cats_vec) <- sankey_cats$names

QoL_sankey <- QoL %>%
  dplyr::select(time, ltfu, sf12_phys_bin, sf12_ment_bin, smwt_dist_bin, stst_nr_bin, phq9_score_bin) %>%
  reshape2::melt(c("time", "ltfu")) %>%
  rename(outcome = variable) %>%
  mutate(outcome = recode(gsub("_bin", "", outcome), !!! outcome_cont),
         outcome = factor(outcome, levels = outcome_cont),
         value = ifelse(ltfu, "ltfu", ifelse(is.na(value), "unknown", ifelse(value == 1, "yes", "no")))) %>%
  rename(category = value) %>%
  group_by(time, outcome, category) %>%
  summarize(value = n()) %>%
  ungroup() %>%
  mutate(time_short = ifelse(time == "Start treatment", "treat. start", ifelse(time == "End treatment", "treat. end", "post treat.")),
         variable = paste(outcome, category),
         variable = recode(variable, !!! sankey_cats_vec),
         variable = paste0(variable, " at ", time_short, " (N=", value, ")"))
 
# A connection data frame is a list of flows with intensity for each flow
links <- rbind(
  QoL_sankey %>%
    filter(time == "Start treatment") %>%
    dplyr::select(outcome, variable) %>%
    rename(source = variable) %>%
    left_join(QoL_sankey %>%
                filter(time == "End treatment") %>%
                dplyr::select(outcome, value, variable) %>%
                rename(target = variable),
              by = "outcome"),
  QoL_sankey %>%
    filter(time == "End treatment") %>%
    dplyr::select(outcome, variable) %>%
    rename(source = variable) %>%
    left_join(QoL_sankey %>%
                filter(time == "Post treatment") %>%
                dplyr::select(outcome, value, variable) %>%
                rename(target = variable),
              by = "outcome"))
 
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$source), as.character(links$target)) %>% unique()
)
nodes$node_group <- stringi::stri_extract(nodes$name, regex = "\\w*-\\w*")
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
 
# Make the Network
colors <- paste(wes_palette("FrenchDispatch"), collapse = '", "')
colorJS <- paste('d3.scaleOrdinal(["', colors, '"])')
sankey_pl <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", NodeGroup = "node_group",
              colourScale = colorJS,
              fontSize = 10,
              sinksRight=FALSE)

sankey_pl <- htmlwidgets::prependContent(sankey_pl, htmltools::tags$h1("Patient flow through treatment by health condition"))
sankey_pl
```

<!-- ## Modeling -->

<!-- ```{r} -->
<!-- # create stan data -->
<!-- stanDF <- QoL_cc_changes %>% -->
<!--   dplyr::filter(!is.na(change)) %>% -->
<!--   mutate(change_type = ifelse(change == -1, "Worsened", ifelse(change == 0, "Unchanged", "Improved")), -->
<!--          # change_type = ifelse(`Start of treatment` == 0, ifelse(change == 1, "improved", "remained good"), -->
<!--          #                      ifelse(change == -1, "worsened", "remained bad")),   -->
<!--          var_change_type = paste(variable, change_type, sep = ": "), -->
<!--          var_change_type = factor(var_change_type, levels = unique(var_change_type)), -->
<!--          var_change_type_int = as.integer(var_change_type), -->
<!--          record_id = factor(record_id, levels = unique(record_id)), -->
<!--          record_id_int = as.integer(stanDF$record_id), -->
<!--          outcome = recode(gsub("_bin", "", variable), !!! outcome_cont), -->
<!--          outcome = factor(outcome, levels = outcome_cont), -->
<!--          change_type = factor(change_type, levels = c("Worsened", "Unchanged", "Improved")) -->
<!--          # change_type = factor(change_type, levels = c("worsened", "remained bad", "remained good", "improved")) -->
<!--          ) -->

<!-- stanList <- list() -->
<!-- stanList$K <- 4 -->
<!-- stanList$V <- n_distinct(stanDF$var_change_type) -->
<!-- stanList$M <- n_distinct(stanDF$record_id) -->
<!-- stanList$N <- nrow(stanDF) -->
<!-- stanList$w <- stanDF$var_change_type_int -->
<!-- stanList$doc <- stanDF$record_id_int -->
<!-- stanList$alpha <- c(1,1,1, 1) -->
<!-- stanList$beta <- rep(1, stanList$V) -->

<!-- # model -->
<!-- lda <- stan(file = "../models/lda-clustering.stan", data = stanList, seed = 12345, cores = 2) -->

<!-- # cluster distr -->
<!-- cluster_distr <- spread_draws(lda, phi[k,v]) %>% -->
<!--   left_join(stanDF %>%  -->
<!--               rename(v = var_change_type_int) %>% -->
<!--               group_by(var_change_type) %>% -->
<!--               slice(1) %>% -->
<!--               ungroup() %>% -->
<!--               dplyr::select(v, outcome, change_type), -->
<!--             by = "v") -->


<!-- cluster_distr %>% -->
<!--   group_by(k, outcome, change_type) %>% -->
<!--   summarize(phi = mean(phi)) %>% -->
<!--   ungroup() %>% -->
<!--   ggplot(aes(x = change_type, y = phi, fill = change_type)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge(width = 1)) + -->
<!--   scale_fill_brewer(palette = "PiYG") + -->
<!--   #scale_fill_manual(values = c("red", "blue")) + -->
<!--   facet_grid(outcome ~ k) + -->
<!--   theme_custom()  -->


<!-- # create change data -->
<!-- df_cc_change <- df_cc %>% -->
<!--   mutate(phq9_score = 27 - phq9_score) %>% # recode PHQ9 Score to facilitate computation of the log changes -->
<!--   group_by(record_id) %>% -->
<!--   summarise( -->
<!--     across(c(site, age, sex, hiv_test_result, drug_resistant, clindiag, cavity, opacity, highbact), ~ .x[1]), -->
<!--     across(c(sf12_phys, sf12_ment, smwt_dist, phq9_score), ~ log(.x[2] / .x[1])), -->
<!--   ) %>% -->
<!--   mutate( -->
<!--     age = ifelse(age < 25, "24y", "25y"), -->
<!--     age = factor(age, levels = c("25y", "24y")), -->
<!--     across(c(drug_resistant, clindiag, cavity, opacity, highbact), ~ ifelse(.x == 1, "Yes", "No")), -->
<!--     across(c(drug_resistant, clindiag, cavity, opacity, highbact), ~ factor(.x, levels = c("No", "Yes"))), -->
<!--     sex = ifelse(sex == 1, "Male", "Female"), -->
<!--     sex = factor(sex, levels = c("Male", "Female")), -->
<!--     site = factor(site, levels = c("Malawi", "Mosambique", "Zimbabwe", "South Africa", "Zambia")), -->
<!--     hiv_test_result = ifelse(hiv_test_result == 1, "Positive", "Negative"), -->
<!--     hiv_test_result = factor(hiv_test_result, c("Negative", "Positive")) -->
<!--   ) -->
<!-- ``` -->




<!-- ### Model change in outcome -->

<!-- ```{r} -->
<!-- # models -->
<!-- mod_sf12phys_out <- stan_glm(sf12_phys ~ 1, data = df_cc_change, prior = NULL) -->
<!-- mod_sf12phys_out$stan_summary %>% data.frame() %>% mutate_all(function(x) round(x * 100, 0)) %>% dplyr::select(`X50.`, `X2.5.`, `X97.5.`) -->
<!-- mod_sf12ment_out <- stan_glm(sf12_ment ~ 1, data = df_cc_change, prior = NULL) -->
<!-- mod_sf12ment_out$stan_summary %>% data.frame() %>% mutate_all(function(x) round(x * 100, 0)) %>% dplyr::select(`X50.`, `X2.5.`, `X97.5.`) -->
<!-- mod_phq9s_out <- stan_glm(phq9_score ~ 1, data = df_cc_change, prior = NULL) -->
<!-- mod_phq9s_out$stan_summary %>% data.frame() %>% mutate_all(function(x) round(x * 100, 0)) %>% dplyr::select(`X50.`, `X2.5.`, `X97.5.`) -->
<!-- mod_smwtd_out <- stan_glm(smwt_dist ~ 1, data = df_cc_change, prior = NULL) -->
<!-- mod_smwtd_out$stan_summary %>% data.frame() %>% mutate_all(function(x) round(x * 100, 0)) %>% dplyr::select(`X50.`, `X2.5.`, `X97.5.`) -->

<!-- draws_occ <- rbind(spread_draws(mod_sf12phys_out, `(Intercept)`) %>% rename(value = `(Intercept)`) %>% mutate(variable = "QoL-PCS"), -->
<!--                    spread_draws(mod_sf12ment_out, `(Intercept)`) %>% rename(value = `(Intercept)`) %>% mutate(variable = "QoL-MCS"), -->
<!--                    spread_draws(mod_phq9s_out, `(Intercept)`) %>% rename(value = `(Intercept)`) %>% mutate(variable = "PHQ9-S"), -->
<!--                    spread_draws(mod_smwtd_out, `(Intercept)`) %>% rename(value = `(Intercept)`) %>% mutate(variable = "6MWT-D")) %>% -->
<!--   mutate(variable = factor(variable, levels = c("QoL-MCS", "QoL-PCS", "6MWT-D", "PHQ9-S")), -->
<!--          value = ifelse(variable == "PHQ9-S", (-1) * value, value)) -->

<!-- sample_sizes <- data.frame(variable = c("QoL-MCS", "QoL-PCS", "6MWT-D", "PHQ9-S"), -->
<!--                            value = .6, -->
<!--                            label = sapply(dplyr::select(df_cc_change, sf12_phys, sf12_ment, phq9_score, smwt_dist), function(x) nrow(df_cc_change) - sum(is.na(x)))) %>% -->
<!--   mutate(label = paste0("N=", label)) -->

<!-- # plot -->
<!-- plot_a <- draws_occ %>% -->
<!--   ggplot(aes(x = variable, y = value)) + -->
<!--   stat_eye(mapping = aes(fill = variable), shape = 23, size = 1, .width = c(.95)) + -->
<!--   #geom_violin(mapping = aes(fill = variable), draw_quantiles = c(0.05, 0.25, 0.5, 0.75, 0.95)) + -->
<!--   #geom_boxplot(width = 0.15, outlier.shape = NA) + -->
<!--   #geom_text(data = sample_sizes, mapping = aes(label = label), size = 8 / cm(1), vjust = -1.5) + -->
<!--   geom_hline(aes(yintercept = 0), linetype = "dotted", color = "red") + -->
<!--   scale_fill_manual(values = wes_palette("Darjeeling2")) + -->
<!--   scale_y_continuous(labels = scales::percent_format(suffix = ""), limits = c(NA, NA), breaks = seq(-1, 1, .25), expand = expansion(add = c(0, .01))) + -->
<!--   coord_flip() + -->
<!--   labs(title = "A: Estimated change in\nhealth scores.", -->
<!--        y = "Estimated change (%)") + -->
<!--   theme_custom() + -->
<!--   theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(),  -->
<!--         legend.position = "top", legend.title = element_blank(), -->
<!--         plot.title = element_text(face = 1, size = 8), -->
<!--         legend.key.width = unit(.4, "cm"), -->
<!--         legend.key.height = unit(.2, "cm"), -->
<!--         panel.spacing = unit(1, "cm"), -->
<!--         strip.text.x = element_text(face = "bold"))  -->

<!-- plot_a -->
<!-- ``` -->

<!-- ### Association with baseline characteristics -->

<!-- ```{r} -->
<!-- # model -->
<!-- mod_sf12ment <- stan_lmer(sf12_ment ~ (1 | site) + age + sex + hiv_test_result + clindiag + cavity,  -->
<!--                           data = df_cc_change, prior = NULL) -->
<!-- mod_sf12phys <- stan_lmer(sf12_phys ~ (1 | site) + age + sex + hiv_test_result + clindiag + cavity,  -->
<!--                           data = df_cc_change, prior = NULL) -->
<!-- mod_smwtd <- stan_lmer(smwt_dist ~ (1 | site) + age + sex + hiv_test_result + clindiag + cavity,  -->
<!--                        data = df_cc_change, prior = NULL) -->
<!-- mod_phq9s <- stan_lmer(phq9_score ~ (1 | site) + age + sex + hiv_test_result + clindiag + cavity,  -->
<!--                        data = df_cc_change, prior = NULL) -->

<!-- # parameters -->
<!-- draws_cc_assoc <- rbind(spread_draws(mod_sf12phys, `sexFemale`, `age24y`, `hiv_test_resultPositive`,  -->
<!--                                      cavityYes, clindiagYes) %>% mutate(outcome = "QoL-PCS"), -->
<!--                         spread_draws(mod_sf12ment, `sexFemale`, `age24y`, `hiv_test_resultPositive`,  -->
<!--                                      cavityYes, clindiagYes) %>% mutate(outcome = "QoL-MCS"), -->
<!--                         spread_draws(mod_phq9s, `sexFemale`, `age24y`, `hiv_test_resultPositive`,  -->
<!--                                      cavityYes, clindiagYes) %>% mutate(outcome = "PHQ9-S"), -->
<!--                         spread_draws(mod_smwtd, `sexFemale`, `age24y`, `hiv_test_resultPositive`,  -->
<!--                                      cavityYes, clindiagYes) %>% mutate(outcome = "6MWT-D")) %>% -->
<!--   rename(`Female` = `sexFemale`, -->
<!--          `Age<25y` = `age24y`, -->
<!--          `HIV+` = hiv_test_resultPositive, -->
<!--          `Cavitation` = `cavityYes`, -->
<!--          `Clin. diagn.` = `clindiagYes`) %>% -->
<!--   dplyr::select(-`.chain`,-`.iteration`,-`.draw`) %>% -->
<!--   melt("outcome") %>% -->
<!--   mutate(outcome = factor(outcome, levels = c("QoL-PCS", "QoL-MCS", "PHQ9-S", "6MWT-D")), -->
<!--          variable = factor(variable, levels = c("Age<25y", "Female", "HIV+", "Clin. diagn.", "Cavitation", "High bact. load")))  -->

<!-- # plot -->
<!-- plot_b <- draws_cc_assoc %>% -->
<!--   group_by(variable, outcome) %>% -->
<!--   median_qi() %>% -->
<!--   ungroup() %>% -->
<!--   ggplot(aes(x = value, xmin = .lower, xmax = .upper, y = variable, color = outcome, group = outcome)) + -->
<!--   geom_errorbar(position = position_dodge(width = .75), width = .5, linewidth = .66) + -->
<!--   geom_pointrange(position = position_dodge(width = .75), shape = 23, fill = "white", linewidth = .66, size = .33) + -->
<!--   #stat_pointinterval(fill = "white", shape = 21, point_size = 2, position = position_dodge2(width = 10), .width = c(0.95)) + -->
<!--   geom_vline(aes(xintercept = 0), linetype = "dotted", color = "red") + -->
<!--   scale_color_manual(values = wes_palette("Darjeeling2")) + -->
<!--   labs(title = "B: Association of changes in health\nscores with baseline characteristics.", -->
<!--        x = "Estimated difference (pp)") + -->
<!--   scale_x_continuous(labels = scales::percent_format(suffix = ""), breaks = seq(-1, 1, .25)) + -->
<!--   theme_custom() +  -->
<!--   theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), -->
<!--         legend.position = "none", -->
<!--         plot.title.position = "plot", -->
<!--         plot.title = element_text(face = 1, size = 8))  -->

<!-- plot_b -->

<!-- legend_grob <- ggplot() + -->
<!--   theme( -->
<!--     panel.background = element_rect(fill='transparent'), #transparent panel bg -->
<!--     plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg -->
<!--     panel.grid.major = element_blank(), #remove major gridlines -->
<!--     panel.grid.minor = element_blank(), #remove minor gridlines -->
<!--     legend.background = element_rect(fill='transparent'), #transparent legend bg -->
<!--     legend.box.background = element_rect(fill='transparent') #transparent legend panel -->
<!--   ) + cowplot::get_legend(plot_a)  -->
<!-- plot_a_nL <- plot_a + theme(legend.position = "none") -->
<!-- plot_b_nL <- plot_b + theme(legend.position = "none") -->
<!-- plot_title <- grid::textGrob("Figure.", gp = grid::gpar(fontsize = 8, fontface = "bold")) -->
<!-- plot_ab <- ((plot_a_nL + plot_spacer() + plot_b_nL) + ((plot_spacer() +  plot_spacer() + legend_grob))) +  -->
<!--   plot_annotation(title = "Figure.", -->
<!--                   caption = "Reference: male, <25y, HIV–, clinically undiagnosed, no cavitation.\nViolin plots: posterior distribution; medians: dots, lines: 95%-CIs. PP=percentage points. ", -->
<!--                   theme = theme_custom()) + -->
<!--   plot_layout(widths = c(1.25, .01, 1.05), heights = c(0.95, 0.05), nrow = 2)  -->
<!-- #save_plot(plot_ab, pdf_file = "../results/croi-abstract-result-plot.pdf", w = 10.15, h = 10.15) -->
<!-- save_plot(plot_ab, pdf_file = "../results/croi-abstract-result-plot.png", w = 10.15, h = 10.15) -->
<!-- ``` -->
